---
title: "EDA - Aki"
author: "Luming Xu, Sijia Zhang, Aki Di Sandro, Ray Ma, Yixuan Zhou"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
    toc_float: true
---

```{r r-setup, results='hide', echo = F, eval = T, warning = F, message = F}
knitr::opts_chunk$set(echo = T, eval = T, warning = F, message = F)

# set working directory
setwd("~/Documents/MUSA/Spring25/MUSA8010_Practicum/Chinatown/")

# load all packages
library(tidyverse)
library(tidycensus)
library(sf)
library(ggplot2)
library(kableExtra)
library(corrplot)
library(googleway)
library(RgoogleMaps)

# read in API
census_api_key("f2855a6037284cb9cbed55e96e6b99be17ee05c6", overwrite = TRUE, install = T)

# mapbox API
# options(rdeck.mapbox_access_token = "pk.eyJ1IjoiY2hpYmlha2kiLCJhIjoiY20xODh2NTNqMTBvaDJqb2ptbjM4ZGViayJ9.un9M1_-S6kI8M0ktqZLz_Q")

# google API
google_api <- colnames(read.delim("~/Documents/MUSA/Spring25/MUSA8010_Practicum/Chinatown/google_api_aki.txt"))

```

# Loading all data

```{r load_data}
# study area
# studyarea <- st_read("data/StudyArea.shp") %>% 
studyarea <- st_read("~/Documents/MUSA/Spring25/MUSA8010_Practicum/Chinatown/Dataset/studyarea/StudyArea.shp") %>%
  st_transform('EPSG:2272')

# Philly block groups
# blockgroups <- st_read("data/Philly_blockgroup.shp") %>%
blockgroups <- st_read("~/Documents/MUSA/Spring25/MUSA8010_Practicum/Chinatown/Dataset/Philly_blockgroup/Philly_blockgroup.shp") %>%
  st_transform('EPSG:2272')

# philly bounds
philly_bounds <- st_union(blockgroups)

# philly hydrology (bounded by philly_bounds; source: https://opendataphilly.org/datasets/hydrology/)
hydro <- st_read("https://services.arcgis.com/fLeGjb7u4uXqeF9q/arcgis/rest/services/Hydrographic_Features_Poly/FeatureServer/1/query?outFields=*&where=1%3D1&f=geojson") %>% 
    st_transform(crs = 'EPSG:2272') %>% 
  st_intersection(philly_bounds)

# highway
# stateroads_inphilly <- st_read("data/PaStateRoads2024_03.geojson") %>% 
#   st_transform('EPSG:2272') %>% 
#   st_intersection(philly_bounds)

# write out stateroads_inphilly as its own file since it's big
# st_write(stateroads_inphilly,
#          "data/PhillyStateRoads.shp")

# read in pre-filte#1a9988 stateroads data
stateroads_inphilly <- st_read("~/Documents/MUSA/Spring25/MUSA8010_Practicum/Chinatown/data/PhillyStateRoads.shp")

philly_mainhighways <- stateroads_inphilly %>%
  filter(TRAF_RT_NO %in% c("I", "US"),
         ST_RT_NO %in% c("0001", "0095", "0076", "0676")) %>%
  dplyr::select(STREET_NAM, ST_RT_NO, TRAF_RT_NO, TRAF_RT__1)

# save and only use the highways of interest
# for now, interested in comparing highways that cut through neighborhoods vs those that don't

# ACS 


# philly property data (https://opendataphilly.org/datasets/philadelphia-properties-and-assessment-history/) <-- that was the original source, but now i'm using the same downloaded file that Luming and Sijia have been working from
philly_properties <- st_read("~/Documents/MUSA/Spring25/MUSA8010_Practicum/Chinatown/data/opa_properties_public.geojson") %>% 
  st_transform(crs = 'EPSG:2272')

# philly park data (https://opendataphilly.org/datasets/ppr-properties/)
philly_parks <- st_read("https://opendata.arcgis.com/datasets/d52445160ab14380a673e5849203eb64_0.geojson") %>%
  st_transform(crs = 'EPSG:2272') %>% 
  st_intersection(philly_bounds)

# philly schools
school <-
  st_read("https://opendata.arcgis.com/datasets/d46a7e59e2c246c891fbee778759717e_0.geojson") %>%
  st_transform('EPSG:2272')

# city facilities (https://opendataphilly.org/datasets/city-facilities-master-facilities-database/)
facilities <- st_read("https://opendata.arcgis.com/datasets/b3c133c3b15d4c96bcd4d5cc09f19f4e_0.geojson") %>%
  st_transform('EPSG:2272') %>% 
  filter(STATUS == "A") # remove inactive sites

# libraries (from facilities data)
libraries <- facilities %>% 
  filter(ASSET_GROUP1 == "A8")

# prisons (from facilities data)
prisons <- facilities %>% 
  filter(ASSET_GROUP2 == "A13.3")

# historic site (from facilities data)
historic <- facilities %>% 
  filter(ASSET_SUBTYPE1 == "C21")

# museums (from facilities data)
museums <- facilities %>% 
  filter(ASSET_GROUP1 == "A18",
         ASSET_SUBTYPE1 == "C35")

# produce (LPSS, HPSS -- indicators for produce; low or high produce supply stores)
retail_produce <- st_read("https://opendata.arcgis.com/datasets/53b8a1c653a74c92b2de23a5d7bf04a0_0.geojson") %>%
  st_transform('EPSG:2272')

# transit stops
metro <- st_read("https://opendata.arcgis.com/api/v3/datasets/af52d74b872045d0abb4a6bbbb249453_0/downloads/data?format=geojson&spatialRefId=4326") %>%
  st_transform('EPSG:2272') %>%
  mutate(Type = "metro")

city_hall <- metro[metro$Station == 'City Hall', 6]

trolley <- st_read("https://opendata.arcgis.com/api/v3/datasets/dd2afb618d804100867dfe0669383159_0/downloads/data?format=geojson&spatialRefId=4326") %>%
  st_transform('EPSG:2272')

# why are some trolley stations missing? going southwest from 40th st station + going northwest from 33rd/34th station

trolley_renamed <- trolley %>%
    rename(Station = StopName,
           Route = LineAbbr,
           Longitude = Lon,
           Latitude = Lat) %>%
    mutate(Type = "trolley")

# Combine both datasets into one
transit <- bind_rows(metro, trolley_renamed)

# PA hospitals (within 2.5 miles of philly bounds)
hospitals <- st_read("~/Documents/MUSA/Spring25/MUSA8010_Practicum/Chinatown/data/DOH_Hospitals202311.geojson") %>% 
  st_transform('EPSG:2272') %>% 
  st_intersection(st_buffer(philly_bounds, 13200))

# checking out what Luming_property.csv looks like
luming_property <- read.csv("~/Documents/MUSA/Spring25/MUSA8010_Practicum/Chinatown/Dataset/Luming_property.csv")
# looking at which properties were saved in luming_property -- Luming and Sijia's properties files are limited to the case study area
# luming_properties <- philly_properties %>% 
#   filter(objectid %in% luming_property$objectid)

# checking out what property_sijia_eda.geojson looks like
sijia_eda <- st_read("~/Documents/MUSA/Spring25/MUSA8010_Practicum/Chinatown/Dataset/property_sijia_eda.geojson")

# google API to get restaurant data
# incomplete, will figure this out later
# restaurant_results <- google_places(search_string = paste(search_terms, city, sep = " "), 
#                                     key = google_api)


```

# Initial plots

```{r plots_initial}
# interstate highways in philly
# ggplot() +
#     geom_sf(data = philly_bounds) +
#     geom_sf(data = stateroads_inphilly %>% 
#               filter(TRAF_RT_NO == "I"))
# 
# # US highways in philly
# ggplot() +
#     geom_sf(data = philly_bounds) +
#     geom_sf(data = stateroads_inphilly %>% 
#               filter(TRAF_RT_NO == "US"))
# 
# # PA state highways in philly
# ggplot() +
#     geom_sf(data = philly_bounds) +
#     geom_sf(data = stateroads_inphilly %>% 
#               filter(TRAF_RT_NO == "PA"))
# 
# # all other roads in philly
# ggplot() +
#     geom_sf(data = philly_bounds) +
#     geom_sf(data = stateroads_inphilly %>% 
#               filter(!(TRAF_RT_NO %in% c("I", "US", "PA"))))

# I-95, US-1, I-676, and I-76
philly_highways_plot <- 
  ggplot() +
  geom_sf(data = philly_bounds, fill = "#c9b887", color = "#57470a") +
  geom_sf(data = hydro, fill = "#7d97c7", color = "transparent") +
  geom_sf(data = philly_parks, fill = "#6b9169", color = "transparent") +
  geom_sf(data = philly_mainhighways,
          aes(color = ST_RT_NO)) +
  # add custom colors
  scale_color_manual(values = c("0001" = "#1a9988",
                                "0076" = "#eb5600",
                                "0095" = "#eba075",
                                "0676" = "#7fe3d6"),
                     labels = c("US-1", "I-76", "I-95", "I-676")) +
  labs(title = "Case Study of Highways in Philly",
       subtitle = "US-1, I-76, I-95, and I-676",
       color = "Highways") +
  theme_void()
philly_highways_plot

# ggsave("Output/HighwaysMap.png", plot = philly_highways_plot)

# looking at study area within philly bounds
ggplot() +
  geom_sf(data = philly_bounds, fill = "#c9b887") + 
  geom_sf(data = hydro, fill = "#7d97c7", color = "transparent") +
  geom_sf(data = studyarea, fill = "coral")

# philly parks
# ggplot() +
#   geom_sf(data = philly_bounds, fill = "#c9b887") + 
#   geom_sf(data = hydro, fill = "#7d97c7", color = "transparent") +
#   geom_sf(data = philly_parks, fill = "#6b9169", color = "transparent")

# philly schools
ggplot() +
  geom_sf(data = philly_bounds, fill = "#c9b887") + 
  geom_sf(data = hydro, fill = "#7d97c7", color = "transparent") +
  geom_sf(data = school, color = "brown4") # point data

# libraries
ggplot() +
  geom_sf(data = philly_bounds, fill = "#c9b887") + 
  geom_sf(data = hydro, fill = "#7d97c7", color = "transparent") +
  geom_sf(data = libraries, color = "#eb5600") # point data

# transit / looking at what this category actually consists of....
# looks like there's point geography for parking lots, some random transit stops, and heli pad station
ggplot() +
  geom_sf(data = philly_bounds, fill = "#c9b887") + 
  geom_sf(data = hydro, fill = "#7d97c7", color = "transparent") +
  geom_sf(data = facilities %>% 
            filter(ASSET_GROUP1 == "A17",
                   grepl("Parking", ASSET_SUBT1_DESC)), color = "pink") # point data

ggplot() +
  geom_sf(data = philly_bounds, fill = "#c9b887") + 
  geom_sf(data = hydro, fill = "#7d97c7", color = "transparent") +
  geom_sf(data = sijia_eda, color = "#7fe3d6")

# prison facilities
ggplot() +
  geom_sf(data = philly_bounds, fill = "#c9b887") + 
  geom_sf(data = hydro, fill = "#7d97c7", color = "transparent") +
  geom_sf(data = facilities %>% 
            filter(ASSET_GROUP2 == "A13.3"), color = "lightgrey") # point data

# all philly properties
# ggplot() +
#   geom_sf(data = philly_bounds, fill = "#c9b887") + 
#   geom_sf(data = hydro, fill = "#7d97c7", color = "transparent") +
#   geom_sf(data = philly_properties, color = "black")

```

# Case Study of Philadelphia's Highways and Housing Price

For all Philly properties, calculate the distance from highway and see how it relates to price.

```{r calc_dist_func}
# taking Luming's distance calculator function
calculate_nearest_distance <- function(set_points, other_layer) {
  nearest_idx <- st_nearest_feature(set_points, other_layer)
  st_distance(set_points, other_layer[nearest_idx, ], by_element = TRUE) %>% as.numeric()
}

```

The following code chunk took too much to load (too many properties to consider), so I'm going to limit it to the properties within 500m (and possibly look at 1000m and 1500m) from highways to look at the immediate effects of highways on property prices.

```{r highwaybuffer}
# only keep properties that are within 500m (1640.42 ft) of highways of interest
buff500 <- 500 * 3.28084
highways_500buffer <- st_union(st_buffer(philly_mainhighways, buff500))

properties500 <- philly_properties[highways_500buffer,]

# only keep properties that are within 1000m (3280.84 ft) of highways of interest
buff1000 <- 1000 * 3.28084
highways_1000buffer <- st_union(st_buffer(philly_mainhighways, buff1000))

properties1000 <- philly_properties[highways_1000buffer,]

# only keep properties that are within 1500m (4921.26 ft) of highways of interest
buff1500 <- 1500 * 3.28084
highways_1500buffer <- st_union(st_buffer(philly_mainhighways, buff1500))

properties1500 <- philly_properties[highways_1500buffer,]

```

## Properties within 500m of highways

```{r temp_plots}
# visualizing buffer
ggplot() +
  geom_sf(data = philly_bounds, fill = "#c9b887", color = "#57470a") +
  geom_sf(data = hydro, fill = "#7d97c7", color = "transparent") +
  geom_sf(data = highways_500buffer,
          color = "transparent", fill = "grey", alpha = 0.4) +
  geom_sf(data = philly_mainhighways,
          aes(color = ST_RT_NO)) +
  
  # add custom colors
  scale_color_manual(values = c("0001" = "#1a9988",
                                "0076" = "#eb5600",
                                "0095" = "#eba075",
                                "0676" = "#7fe3d6")) +
  
  labs(title = "Case Study of Highways in Philly",
       subtitle = "US-1, I-76, I-95, and I-676",
       color = "Highways",
       fill = "Highways") +
  theme_void()

# visualizing properties to make sure we got the right ones
ggplot() +
  geom_sf(data = philly_bounds, fill = "#c9b887", color = "#57470a") +
  geom_sf(data = hydro, fill = "#7d97c7", color = "transparent") +
  geom_sf(data = philly_parks, fill = "#6b9169", color = "transparent") +
  geom_sf(data = highways_500buffer,
          color = "transparent", fill = "grey", alpha = 0.4) +
  geom_sf(data = properties500,
          color = "#4a4a4a", size = .3) +
  geom_sf(data = philly_mainhighways,
          aes(color = ST_RT_NO)) +
  
  # add custom colors
  scale_color_manual(values = c("0001" = "#1a9988",
                                "0076" = "#eb5600",
                                "0095" = "#eba075",
                                "0676" = "#7fe3d6")) +
  
  labs(title = "Properties within 500m of Highways in Philly",
       subtitle = "US-1, I-76, I-95, and I-676",
       color = "Highways",
       fill = "Highways") +
  theme_void()

ggplot() +
  geom_sf(data = philly_bounds, fill = "#c9b887", color = "#57470a") +
  geom_sf(data = hydro, fill = "#7d97c7", color = "transparent") +
  geom_sf(data = philly_parks, fill = "#6b9169", color = "transparent") +
  geom_sf(data = highways_1000buffer,
          color = "transparent", fill = "grey", alpha = 0.4) +
  geom_sf(data = properties1000,
          color = "#4a4a4a", size = .3) +
  geom_sf(data = philly_mainhighways,
          aes(color = ST_RT_NO)) +
  
  # add custom colors
  scale_color_manual(values = c("0001" = "#1a9988",
                                "0076" = "#eb5600",
                                "0095" = "#eba075",
                                "0676" = "#7fe3d6"),
                       labels = c("US-1", "I-76", "I-95", "I-676")) +
  
  labs(title = "Properties within 1000m of Highways in Philly",
       subtitle = "US-1, I-76, I-95, and I-676",
       color = "Highways",
       fill = "Highways") +
  theme_void()

ggplot() +
  geom_sf(data = philly_bounds, fill = "#c9b887", color = "#57470a") +
  geom_sf(data = hydro, fill = "#7d97c7", color = "transparent") +
  geom_sf(data = philly_parks, fill = "#6b9169", color = "transparent") +
  geom_sf(data = highways_1500buffer,
          color = "transparent", fill = "grey", alpha = 0.4) +
  geom_sf(data = properties1500,
          color = "#4a4a4a", size = .3) +
  geom_sf(data = philly_mainhighways,
          aes(color = ST_RT_NO)) +
  
  # add custom colors
  scale_color_manual(values = c("0001" = "#1a9988",
                                "0076" = "#eb5600",
                                "0095" = "#eba075",
                                "0676" = "#7fe3d6")) +
  
  labs(title = "Properties within 1500m of Highways in Philly",
       subtitle = "US-1, I-76, I-95, and I-676",
       color = "Highways",
       fill = "Highways") +
  theme_void()

```

Are there differences between how price of property relates to dist_to_closest_highway depending on type of highway (whether highway cuts through neighborhood or not)?

```{r clean_prop500}
# starting with 500m buffer
# need to check if there are 0 values in sale_price before log transforming
summary(properties500$sale_price)
# Min.   1st Qu.    Median      Mean   3rd Qu.      Max.      NA's 
#   0        10     89900    327917    230000 342000000       840 

# only keep properties whose sale_price >= $10,000
# according to #1a9988fin, the most expensive house in philadelphia right now is $25,000,000 -- using that as a cutoff
properties500_clean <- properties500 %>% 
  filter(sale_price >= 10000,
         sale_price <= 25000000,
         total_area >= 100,
         total_livable_area >= 100,
         sale_date >= as.Date("2005-01-01", format = "%Y-%m-%d"),
         sale_date <= Sys.Date()) %>% 
  mutate(price_perTLA = sale_price / total_livable_area,
         price_perTA = sale_price / total_area,
         log_price = log(sale_price),
         log_price_perTLA = log(price_perTLA),
         log_price_perTA = log(price_perTA),
         norm_log_price = scale(log_price),
         norm_log_price_perTLA = scale(log_price_perTLA),
         norm_log_price_perTA = scale(log_price_perTA),
         category_easy = case_when(category_code_description %in% c("APARTMENTS  > 4 UNITS",
                                                                    "MIXED USE",
                                                                    "MULTI FAMILY",
                                                                    "SINGLE FAMILY") ~ "Residential or Mixed",
                                   grepl("INDUS", category_code_description) ~ "Industrial",
                                   grepl("HOTEL", category_code_description) ~ "Hotel",
                                   grepl("OFFICES", category_code_description) ~ "Offices",
                                   category_code_description %in% c("COMMERCIAL",
                                                                    "GARAGE - COMMERCIAL",
                                                                    "RETAIL") ~ "Commercial",
                                   grepl("VACANT", category_code_description) ~ "Vacant",
                                   .default = "Other"))

summary(properties500_clean$sale_price)
#  Min.  1st Qu.   Median     Mean  3rd Qu.     Max. 
# 10000   103000   174900   340263   295000 24750000
summary(properties500_clean$price_perTLA)
# Min.  1st Qu.   Median     Mean  3rd Qu.     Max. 
# 0.77    78.86   130.36   181.26   208.33 11527.5
summary(properties500_clean$price_perTA)
# Min.  1st Qu.   Median     Mean  3rd Qu.     Max. 
# 0.23    58.35    99.21   218.95   220.79 22671.57

# correlations between total_area and total_livable_area, and sale_price
# cor(properties500_clean$total_area, properties500_clean$total_livable_area, use = "pairwise.complete.obs") # 0.5443428
# cor(properties500_clean$total_area, properties500_clean$sale_price, use = "pairwise.complete.obs") # 0.4210804
# cor(properties500_clean$total_livable_area, properties500_clean$sale_price, use = "pairwise.complete.obs") # 0.577406

# histogram of sale_price
ggplot(data = properties500_clean) +
  geom_histogram(aes(x = sale_price),
                 bins = 100)

ggplot(data = properties500_clean) +
  geom_histogram(aes(x = price_perTA),
                 bins = 100)

ggplot(data = properties500_clean) +
  geom_histogram(aes(x = price_perTLA),
                 bins = 100)

ggplot(data = properties500_clean) +
  geom_histogram(aes(x = log_price),
                 bins = 100)

ggplot(data = properties500_clean) +
  geom_histogram(aes(x = log_price_perTA),
                 bins = 100)

ggplot(data = properties500_clean) +
  geom_histogram(aes(x = log_price_perTLA),
                 bins = 100)

```


```{r maps_prop500}
# map of properties within 500 m of highway, colored by sale_price
# ggplot() +
#   geom_sf(data = philly_bounds, fill = "#c9b887", color = "grey") +
#   geom_sf(data = hydro, fill = "#7d97c7", color = "transparent") +
#   geom_sf(data = philly_parks, fill = "#6b9169", color = "transparent") +
#   geom_sf(data = highways_500buffer,
#           color = "transparent", fill = "grey", alpha = 0.4) +
#   geom_sf(data = properties500_clean,
#           aes(color = sale_price), size = .3) +
#   geom_sf(data = philly_mainhighways,
#           color = "black") +
#   labs(title = "Sale Price of Properties within 500m of Highways in Philly",
#        subtitle = "US-1, I-76, I-95, and I-676") +
#   theme_void()

# map of properties within 500 m of highway, colored by price_perTA
# ggplot() +
#   geom_sf(data = philly_bounds, fill = "#c9b887", color = "grey") +
#   geom_sf(data = hydro, fill = "#7d97c7", color = "transparent") +
#   geom_sf(data = philly_parks, fill = "#6b9169", color = "transparent") +
#   geom_sf(data = highways_500buffer,
#           color = "transparent", fill = "grey", alpha = 0.4) +
#   geom_sf(data = properties500_clean,
#           aes(color = price_perTA), size = .3) +
#   geom_sf(data = philly_mainhighways,
#           color = "black") +
#   labs(title = "Sale Price (per Total Area) of Properties within 500m of Highways in Philly",
#        subtitle = "US-1, I-76, I-95, and I-676") +
#   theme_void()

# ggplot() +
#   geom_sf(data = philly_bounds, fill = "#c9b887", color = "grey") +
#   geom_sf(data = hydro, fill = "#7d97c7", color = "transparent") +
#   geom_sf(data = philly_parks, fill = "#6b9169", color = "transparent") +
#   geom_sf(data = highways_500buffer,
#           color = "transparent", fill = "grey", alpha = 0.4) +
#   geom_sf(data = properties500_clean,
#           aes(color = price_perTLA), size = .3) +
#   geom_sf(data = philly_mainhighways,
#           color = "black") +
#   labs(title = "Sale Price (per Total Livable Area) of Properties within 500m of Highways in Philly",
#        subtitle = "US-1, I-76, I-95, and I-676") +
#   theme_void()

ggplot() +
  geom_sf(data = philly_bounds, fill = "#c9b887", color = "grey") +
  geom_sf(data = hydro, fill = "#7d97c7", color = "transparent") +
  geom_sf(data = philly_parks, fill = "#6b9169", color = "transparent") +
  geom_sf(data = highways_500buffer,
          color = "transparent", fill = "black", alpha = 0.2) +
  geom_sf(data = properties500_clean,
          aes(color = log_price), size = .3) +
  geom_sf(data = philly_mainhighways,
          color = "black") +
  labs(title = "Log Sale Price of Properties within 500m of Highways in Philly",
       subtitle = "US-1, I-76, I-95, and I-676",
       color = "Log Price") +
  theme_void()

# house sale price (category_easy == "Residential or Mixed")
ggplot() +
  geom_sf(data = philly_bounds, fill = "#c9b887", color = "grey") +
  geom_sf(data = hydro, fill = "#7d97c7", color = "transparent") +
  geom_sf(data = philly_parks, fill = "#6b9169", color = "transparent") +
  geom_sf(data = highways_500buffer,
          color = "transparent", fill = "black", alpha = 0.2) +
  geom_sf(data = properties500_clean %>% 
            filter(grepl("Resident", category_easy)),
          aes(color = log_price), size = .3) +
  geom_sf(data = philly_mainhighways,
          color = "black") +
  labs(title = "Log Sale Price of Properties within 500m of Highways in Philly",
       subtitle = "Residential or Mixed Use properties only\nUS-1, I-76, I-95, and I-676",
       color = "Log Price") +
  theme_void()

ggplot() +
  geom_sf(data = philly_bounds, fill = "#c9b887", color = "grey") +
  geom_sf(data = hydro, fill = "#7d97c7", color = "transparent") +
  geom_sf(data = philly_parks, fill = "#6b9169", color = "transparent") +
  geom_sf(data = highways_500buffer,
          color = "transparent", fill = "grey", alpha = 0.4) +
  geom_sf(data = properties500_clean,
          aes(color = log_price_perTA), size = .3) +
  geom_sf(data = philly_mainhighways,
          color = "black") +
  labs(title = "Log Sale Price (per Total Area) of Properties within 500m of Highways in Philly",
       subtitle = "US-1, I-76, I-95, and I-676",
       color = "Log Price (per TA)") +
  theme_void()

# same thing but residential or mixed use only
ggplot() +
  geom_sf(data = philly_bounds, fill = "#c9b887", color = "grey") +
  geom_sf(data = hydro, fill = "#7d97c7", color = "transparent") +
  geom_sf(data = philly_parks, fill = "#6b9169", color = "transparent") +
  geom_sf(data = highways_500buffer,
          color = "transparent", fill = "grey", alpha = 0.4) +
  geom_sf(data = properties500_clean %>% 
            filter(grepl("Resident", category_easy)),
          aes(color = log_price_perTA), size = .3) +
  geom_sf(data = philly_mainhighways,
          color = "black") +
  labs(title = "Log Sale Price (per Total Area) of Properties within 500m of Highways in Philly",
       subtitle = "Residential or Mixed Use properties only\nUS-1, I-76, I-95, and I-676",
       color = "Log Price (per TA)") +
  theme_void()

ggplot() +
  geom_sf(data = philly_bounds, fill = "#c9b887", color = "grey") +
  geom_sf(data = hydro, fill = "#7d97c7", color = "transparent") +
  geom_sf(data = philly_parks, fill = "#6b9169", color = "transparent") +
  geom_sf(data = highways_500buffer,
          color = "transparent", fill = "grey", alpha = 0.4) +
  geom_sf(data = properties500_clean,
          aes(color = log_price_perTLA), size = .3) +
  geom_sf(data = philly_mainhighways,
          color = "black") +
  labs(title = "Log Sale Price (per Total Livable Area) of Properties within 500m of Highways in Philly",
       subtitle = "US-1, I-76, I-95, and I-676",
       color = "Log Price (per TLA)") +
  theme_void()

# same thing but residential/mixed use only
ggplot() +
  geom_sf(data = philly_bounds, fill = "#c9b887", color = "grey") +
  geom_sf(data = hydro, fill = "#7d97c7", color = "transparent") +
  geom_sf(data = philly_parks, fill = "#6b9169", color = "transparent") +
  geom_sf(data = highways_500buffer,
          color = "transparent", fill = "grey", alpha = 0.4) +
  geom_sf(data = properties500_clean %>% 
            filter(grepl("Resident", category_easy)),
          aes(color = log_price_perTLA), size = .3) +
  geom_sf(data = philly_mainhighways,
          color = "black") +
  labs(title = "Log Sale Price (per Total Livable Area) of Properties within 500m of Highways in Philly",
       subtitle = "Residential or Mixed Use properties only\nUS-1, I-76, I-95, and I-676",
       color = "Log Price (per TLA)") +
  theme_void()

```

Calculate each property's distance to highways.

```{r calc_dist2highway}
# calculate each property's distance to highways of interest
# distance is in feet
# start_time <- Sys.time()
prop500 <- properties500_clean %>% 
  select(matches("assessment_date|building_code|category_|census_tract|geographic|zip|market_value|parcel|sale_|taxable|_area|year_built|objectid|geometry|price")) %>% 
  mutate(dist_to_US001 = calculate_nearest_distance(geometry,
                                     philly_mainhighways %>%
                                       filter(ST_RT_NO == "0001")),
         dist_to_I076 = calculate_nearest_distance(geometry,
                                    philly_mainhighways %>%
                                      filter(ST_RT_NO == "0076")),
         dist_to_I095 = calculate_nearest_distance(geometry,
                                    philly_mainhighways %>%
                                      filter(ST_RT_NO == "0095")),
         dist_to_I676 = calculate_nearest_distance(geometry,
                                    philly_mainhighways %>%
                                      filter(ST_RT_NO == "0676")),
         dist_to_US001m = dist_to_US001 * 0.3048,
         dist_to_I076m = dist_to_I076 * 0.3048,
         dist_to_I095m = dist_to_I095 * 0.3048,
         dist_to_I676m = dist_to_I676 * 0.3048) # Time difference of 55.20697 secs

# find closest highway
prop500_closest <- prop500 %>%
  select(matches("objectid|dist_to")) %>%
  pivot_longer(cols = 2:5,
               names_to = "highway",
               values_to = "distance") %>%
  group_by(objectid) %>%
  summarise(dist_to_closest_highway = min(distance, na.rm = T)) %>%
  ungroup()  %>% 
  mutate(dist_to_closest_highwaym = dist_to_closest_highway * 0.3048) 

# join closest highway
prop500 <- left_join(prop500,
                     prop500_closest %>% st_drop_geometry(),
                     by = "objectid") %>%
  mutate(closest_highway = case_when(
    dist_to_closest_highway == dist_to_US001 ~ "US-1",
    dist_to_closest_highway == dist_to_I076 ~ "I-76",
    dist_to_closest_highway == dist_to_I095 ~ "I-95",
    dist_to_closest_highway == dist_to_I676 ~ "I-676",
    .default = NA),
    closest_highway = factor(closest_highway, levels = c("US-1", "I-76", "I-676", "I-95")),
    highway_type = case_when(closest_highway %in% c("US-1","I-676") ~ "through neighborhood",
                             closest_highway %in% c("I-95","I-76") ~ "along waterway",
                             .default = NA))

# save file
# st_write(prop500, "/Users/akiradisandro/Documents/MUSA/Spring25/MUSA8010_Practicum/Chinatown/data/philly_prop500_dist2highway.geojson")

# read in pre-saved file
# prop500_new <- st_read("/Users/akiradisandro/Documents/MUSA/Spring25/MUSA8010_Practicum/Chinatown/Dataset/data/philly_prop500_dist2highway.geojson")

```

First, let's looks at a faceted plot of the distribution of sales price for each highway separately

```{r prop500_faceted_distr}
# prop500_faceted <- 
#   ggplot(data = prop500,
#          aes(x = dist_to_closest_highwaym, y = log_price_perTA)) +
#   geom_point(color = "#1a9988", alpha = .1, size = 0.3) +
#   geom_smooth(color = "#eb5600", method = "lm", se = F, linewidth = 1.2) +
#   # geom_point(alpha = .1, size = 0.3) +
#   # geom_smooth(method = "lm", se = F, linewidth = 1.2) +
#   facet_wrap(~closest_highway) +
#   labs(title = "Distribution of Property Sale Price",
#        x = "Distance to Closest Highway (m)",
#        y = "Log Sale Price (per TA)") +
#   theme_minimal()

prop500_faceted <- 
  ggplot(data = prop500,
         aes(x = log_price_perTA, color = closest_highway, fill = closest_highway)) +
  geom_histogram(bins = 50) +
  facet_wrap(~closest_highway, scales = "free_y") +
  scale_color_manual(values = c("US-1" = "#1a9988",
                                "I-76" = "#eb5600",
                                "I-95" = "#eba075",
                                "I-676" = "#7fe3d6")) +
  scale_fill_manual(values = c("US-1" = "#1a9988",
                               "I-76" = "#eb5600",
                               "I-95" = "#eba075",
                               "I-676" = "#7fe3d6")) +
  labs(title = "Distribution of Property Sale Price",
       x = "Log Sale Price (per Total Area)",
       y = "Count",
       color = "Closest Highway",
       fill = "Closest Highway") +
  theme_minimal()

prop500_faceted

# save image for slides
# ggsave("Output/SalePriceDistribution.png", plot = prop500_faceted)

# residential + mixed only 
# prop500_faceted_res <- 
#   ggplot(data = prop500 %>% 
#            filter(grepl("Reside", category_easy)),
#          aes(x = dist_to_closest_highwaym, y = log_price_perTA)) +
#   geom_point(color = "#1a9988", alpha = .1, size = 0.3) +
#   geom_smooth(color = "#eb5600", method = "lm", se = F, linewidth = 1.2) +
#   # geom_point(alpha = .1, size = 0.3) +
#   # geom_smooth(method = "lm", se = F, linewidth = 1.2) +
#   # facet_wrap(closest_highway) +
#   labs(title = "Distribution of Property Sale Price",
#        x = "Distance to Closest Highway (m)",
#        y = "Log Sale Price (per TA)") +
#   theme_minimal()

prop500_faceted_res <- 
  ggplot(data = prop500 %>% 
           filter(grepl("Resid", category_easy)),
         aes(x = log_price_perTA, color = closest_highway, fill = closest_highway)) +
  geom_histogram(bins = 50) +
  facet_wrap(~closest_highway, scales = "free_y") +
  scale_color_manual(values = c("US-1" = "#1a9988",
                                "I-76" = "#eb5600",
                                "I-95" = "#eba075",
                                "I-676" = "#7fe3d6")) +
  scale_fill_manual(values = c("US-1" = "#1a9988",
                               "I-76" = "#eb5600",
                               "I-95" = "#eba075",
                               "I-676" = "#7fe3d6")) +
  labs(title = "Distribution of Property Sale Price",
       subtitle = "Residential and Mixed Properties only",
       x = "Log Sale Price (per Total Area)",
       y = "Count",
       color = "Closest Highway",
       fill = "Closest Highway") +
  theme_minimal()

prop500_faceted_res

# save image for slides
# ggsave("Output/SalePriceDistribution_residential.png", plot = prop500_faceted_res)

```

Now, look at relationship between distance to highway and price.

```{r price_dist2highway}
# map of properties within 500m of highway cored by which highway they're closest to
ggplot() +
  geom_sf(data = philly_bounds, fill = "#c9b887", color = "grey") +
  geom_sf(data = hydro, fill = "#7d97c7", color = "transparent") +
  geom_sf(data = philly_parks, fill = "#6b9169", color = "transparent") +
  geom_sf(data = highways_500buffer,
          color = "transparent", fill = "grey", alpha = 0.4) +
  geom_sf(data = prop500,
          aes(color = closest_highway), size = 0.3) +
  geom_sf(data = philly_mainhighways,
          color = "black") +
  
  # add custom colors
  scale_color_manual(values = c("US-1" = "#1a9988",
                                "I-76" = "#eb5600",
                                "I-95" = "#eba075",
                                "I-676" = "#7fe3d6")) +
  
  labs(title = "Properties within 500m of Highways in Philly",
       subtitle = "US-1, I-76, I-95, and I-676",
       color = "Closest Highway") +
  theme_void()

# simple scatter of price and dist2closesthighway
ggplot(data = prop500, 
       aes(x = dist_to_closest_highwaym, y = sale_price, 
           color = closest_highway, fill = closest_highway)) +
  geom_point(alpha = .2) +
  geom_smooth(method = "lm") +
  scale_color_manual(values = c("US-1" = "#1a9988",
                                "I-76" = "#eb5600",
                                "I-95" = "#eba075",
                                "I-676" = "#7fe3d6")) +
  scale_fill_manual(values = c("US-1" = "#1a9988",
                               "I-76" = "#eb5600",
                               "I-95" = "#eba075",
                               "I-676" = "#7fe3d6")) +
  labs(title = "Property Price as a function of Distance to Closest Highway",
       subtitle = "Colored by Closest Highway; Properties within 500m of highways",
       fill = "Closest Highway",
       color = "Closest Highway",
       x = "Distance to Closest Highway (m)",
       y = "Property Sale Price ($)") +
  theme_minimal()

# simple scatter of price (as log price per total area in sq ft) and dist2closesthighway
scatter_4highways <- 
  ggplot(data = prop500, 
       aes(x = dist_to_closest_highwaym, y = log_price_perTA, 
           color = closest_highway, fill = closest_highway)) +
  geom_point(alpha = .2, size = 0.3) +
  geom_smooth(method = "lm", size = 1.2) +
  scale_color_manual(values = c("US-1" = "#1a9988",
                                "I-76" = "#eb5600",
                                "I-95" = "#eba075",
                                "I-676" = "#7fe3d6")) +
  scale_fill_manual(values = c("US-1" = "#1a9988",
                               "I-76" = "#eb5600",
                               "I-95" = "#eba075",
                               "I-676" = "#7fe3d6")) +
  labs(title = "Property Price (log price per TA) as a function of Distance to Closest Highway",
       subtitle = "Colored by Closest Highway; Properties within 500m of highways",
       fill = "Closest Highway",
       color = "Closest Highway",
       x = "Distance to Closest Highway (m)",
       y = "Property Sale Price (log price per TA)") +
  theme_minimal()

scatter_4highways

# save image for slides
# ggsave("Output/PriceVSDist2Highway_4highways.png",
#        plot = scatter_4highways,
#        height = 7,
#        width = 10)

# simple scatter of price (as log price per total area in sq ft) and dist2closesthighway 
# sale_price <= 1e7
ggplot(data = prop500 %>% 
         filter(sale_price <= 1e7), 
       aes(x = dist_to_closest_highwaym, y = log_price_perTA, 
           color = closest_highway, fill = closest_highway)) +
  geom_point(alpha = .2, size = 0.3) +
  geom_smooth(method = "lm", size = 1.2) +
  scale_color_manual(values = c("US-1" = "#1a9988",
                                "I-76" = "#eb5600",
                                "I-95" = "#eba075",
                                "I-676" = "#7fe3d6")) +
  scale_fill_manual(values = c("US-1" = "#1a9988",
                               "I-76" = "#eb5600",
                               "I-95" = "#eba075",
                               "I-676" = "#7fe3d6")) +
  labs(title = "Property Price (log price per TA) as a function of Distance to Closest Highway",
       subtitle = "Colored by Closest Highway; Properties within 500m of highways",
       fill = "Closest Highway",
       color = "Closest Highway",
       x = "Distance to Closest Highway (m)",
       y = "Property Sale Price (log price per TA)") +
  theme_minimal()

# simple scatter of price (as log price per total area in sq ft) and log(dist2closesthighway)
ggplot(data = prop500, 
       aes(x = log(dist_to_closest_highwaym), y = log_price_perTA, 
           color = closest_highway, fill = closest_highway)) +
  geom_point(alpha = .2, size = 0.3) +
  geom_smooth(method = "lm", size = 1.2) +
  scale_color_manual(values = c("US-1" = "#1a9988",
                                "I-76" = "#eb5600",
                                "I-95" = "#eba075",
                                "I-676" = "#7fe3d6")) +
  scale_fill_manual(values = c("US-1" = "#1a9988",
                               "I-76" = "#eb5600",
                               "I-95" = "#eba075",
                               "I-676" = "#7fe3d6")) +
  labs(title = "Property Price (log price per TA) as a function of Distance to Closest Highway (log)",
       subtitle = "Colored by Closest Highway; Properties within 500m of highways",
       fill = "Closest Highway",
       color = "Closest Highway",
       x = "Log Distance to Closest Highway (m)",
       y = "Log Property Sale Price (per TA)") +
  theme_minimal()

```

```{r prop500_regression}
# simple regression of price (log_price_perTA) and dist2closesthighway
fit1 <- lm(log_price_perTA ~ dist_to_closest_highwaym, data = prop500)

summary(fit1)

# multiple regression of price (log_price_perTA) and dist2closesthighway but adding highway type (4 types)
fit2 <- lm(log_price_perTA ~ dist_to_closest_highwaym + closest_highway, data = prop500)

summary(fit2)

# multiple regression of price (log_price_perTA) and dist2closesthighway but adding highway type (4 types) with interaction
fit3 <- lm(log_price_perTA ~ dist_to_closest_highwaym + closest_highway + dist_to_closest_highwaym*closest_highway, data = prop500)

summary(fit3)

# multiple regression of price (log_price_perTA) and dist2closesthighway but adding highway type (2 types)
fit4 <- lm(log_price_perTA ~ dist_to_closest_highwaym + highway_type, data = prop500)

summary(fit4)

# multiple regression of price (log_price_perTA) and dist2closesthighway but adding highway type (2 types) with interaction
fit5 <- lm(log_price_perTA ~ dist_to_closest_highwaym + highway_type + dist_to_closest_highwaym*highway_type, data = prop500)

sum5 <- summary(fit5)

data.frame(sum5$coefficients) %>% 
  mutate(Estimate = round(Estimate, 3),
         StdErorr = round(Std..Error, 3),
         PValue = round(Pr...t.., 3)) %>% 
  select(Estimate, StdErorr, PValue) %>% 
  kbl(align = c("l", rep("c", 3)),
      caption = "Linear Regression: Sale Price ~ Distance to Closest Highway + Highway Type + Interaction") %>% 
    kable_styling() %>% 
    kable_classic(full_width = F, html_font = "Cambria") 


# multiple regression of price (log_price_perTA) and dist2closesthighway but adding highway type (2 types) with interaction
fit6 <- lm(log_price_perTA ~ log(dist_to_closest_highwaym) + highway_type + log(dist_to_closest_highwaym)*highway_type, data = prop500)

summary(fit6)

```


Visualizing the highway types (cutting through neighborhoods vs running alongside water).
```{r price_dist2highway_2types}
# simple scatter of price and dist2closesthighway
ggplot(data = prop500, 
       aes(x = dist_to_closest_highwaym, y = sale_price, 
           color = highway_type, fill = highway_type)) +
  geom_point(alpha = .2, size = 0.3) +
  geom_smooth(method = "lm") +
  scale_color_manual(values = c("through neighborhood" = "#1a9988",
                                "along waterway" = "#eb5600")) +
  scale_fill_manual(values = c("through neighborhood" = "#1a9988",
                               "along waterway" = "#eb5600")) +
  labs(title = "Property Price as a function of Distance to Highway Type",
       subtitle = "Colored by Highway Type; Properties within 500m of highways",
       fill = "Highway Type",
       color = "Highway Type") +
  theme_minimal()

# simple scatter of price and dist2closesthighway with loess method
ggplot(data = prop500, 
       aes(x = dist_to_closest_highwaym, y = sale_price, 
           color = highway_type, fill = highway_type)) +
  geom_point(alpha = .2, size = 0.3) +
  geom_smooth(method = "loess", se = FALSE, size = 1.2) +
  scale_color_manual(values = c("through neighborhood" = "#1a9988",
                                "along waterway" = "#eb5600")) +
  scale_fill_manual(values = c("through neighborhood" = "#1a9988",
                               "along waterway" = "#eb5600")) +
  labs(title = "Property Price as a function of Distance to Highway Type",
       subtitle = "Colored by Highway Type; Properties within 500m of highways",
       fill = "Highway Type",
       color = "Highway Type") +
  theme_minimal()

# simple scatter of price (as log price per total area in sq ft) and dist2closesthighway
scatter_2types <- 
  ggplot(data = prop500, 
       aes(x = dist_to_closest_highwaym, y = log_price_perTA, 
           color = highway_type, fill = highway_type)) +
  geom_point(alpha = .2, size = 0.3) +
  geom_smooth(method = "lm", size = 1.2) +
  scale_color_manual(values = c("through neighborhood" = "#1a9988",
                                "along waterway" = "#eb5600")) +
  scale_fill_manual(values = c("through neighborhood" = "#1a9988",
                               "along waterway" = "#eb5600")) +
  labs(title = "Property Price (log price per TA) as a function of Distance to Highway Type",
       subtitle = "Colored by Highway Type; Properties within 500m of highways",
       fill = "Highway Type",
       color = "Highway Type",
       x = "Distance to Closest Highway (m)",
       y = "Log Sale Price (per TA)") +
  theme_minimal()
scatter_2types

# save for slides
# ggsave("Output/PriceVSDist2Highway_2types.png", 
#        plot = scatter_2types,
#        height = 7,
#        width = 10)

# simple scatter of price (as log price per total area in sq ft) and dist2closesthighway
ggplot(data = prop500, 
       aes(x = log(dist_to_closest_highwaym), y = log_price_perTA, 
           color = highway_type, fill = highway_type)) +
  geom_point(alpha = .2, size = 0.3) +
  geom_smooth(method = "lm", size = 1.2) +
  scale_color_manual(values = c("through neighborhood" = "#1a9988",
                                "along waterway" = "#eb5600")) +
  scale_fill_manual(values = c("through neighborhood" = "#1a9988",
                               "along waterway" = "#eb5600")) +
  labs(title = "Property Price (log price per TA) as a function of Distance to Highway Type",
       subtitle = "Colored by Highway Type; Properties within 500m of highways",
       fill = "Highway Type",
       color = "Highway Type",
       x = "Distance to Closest Highway (m)",
       y = "Log Sale Price (per TA)") +
  theme_minimal()

# correlations
prop500 %>%
  st_drop_geometry() %>%
  group_by(highway_type) %>%
  summarize(correlation = cor(dist_to_closest_highwaym, log_price_perTA, method = "pearson"), .groups = "drop") %>% 
  mutate(correlation = round(correlation, 3)) %>% 
  kbl(col.names = c("Highway Type", "Correlation"),
        align = c("l", "c"),
      caption = "Correlation Between Property Sale Price and Distance to Closest Highway by Highway Types") %>% 
    kable_styling() %>% 
    kable_classic(full_width = F, html_font = "Cambria") 

# correlation of price and dist to highway for each highway
prop500 %>%
  st_drop_geometry() %>%
  group_by(closest_highway) %>%
  summarize(correlation = cor(dist_to_closest_highwaym, log_price_perTA, method = "pearson"), .groups = "drop") %>% 
  mutate(correlation = round(correlation, 3)) %>% 
  kbl(col.names = c("Highway", "Correlation"),
        align = c("l", "c"),
      caption = "Correlation Between Property Sale Price and Distance to Closest Highway by Highway") %>%
  kable_styling() %>% 
  kable_classic(html_font = "Cambria") 

prop500 %>%
  st_drop_geometry() %>%
  group_by(closest_highway) %>%
  summarize(correlation = cor(log(dist_to_closest_highwaym), log_price_perTA, method = "pearson"), .groups = "drop")

```



Limiting the properties further to those under $1mil

```{r quiet = TRUE, results = 'hide'}
prop500_under1m <- prop500 %>%
  filter(sale_price < 1e6)
```

```{r}
ggplot(data = prop500_under1m, 
       aes(x = dist_to_closest_highwaym, y = sale_price, 
           color = closest_highway, fill = closest_highway)) +
  geom_point(alpha = .1, size = 0.3) +
  geom_smooth(method = "loess", se = FALSE, size = 1.2) +
  scale_color_manual(values = c("US-1" = "#1a9988",
                                "I-76" = "#eb5600",
                                "I-95" = "#eba075",
                                "I-676" = "#7fe3d6")) +
  scale_fill_manual(values = c("US-1" = "#1a9988",
                               "I-76" = "#eb5600",
                               "I-95" = "#eba075",
                               "I-676" = "#7fe3d6")) +
  labs(title = "Property Price (under $1M) as a function of Distance to Closest Highway",
       subtitle = "Colored by Closest Highway; Properties within 500m of highways",
       fill = "Closest Highway",
       color = "Closest Highway") +
  theme_minimal()
```

```{r cor_under1M}
# correlation of price and dist to highway for each highway
cor <- prop500_under1m %>%
  st_drop_geometry() %>%
  group_by(closest_highway) %>%
  summarize(correlation = cor(dist_to_closest_highway, log_price_perTA, method = "pearson"), .groups = "drop")
cor

# correlation of price and dist to closest highway
prop500_under1m %>%
  st_drop_geometry() %>%
  summarize(correlation = cor(dist_to_closest_highway, log_price_perTA, method = "pearson"), .groups = "drop")

```

```{r}
# simple scatter of price (as log price per total area in sq ft) and dist2closesthighway
ggplot(data = prop500_under1m, 
       aes(x = dist_to_closest_highwaym, y = log_price_perTA, 
           color = closest_highway, fill = closest_highway)) +
  geom_point(alpha = .1, size = 0.3) +
  geom_smooth(method = "lm", se=F, size = 1.2) +
  scale_color_manual(values = c("US-1" = "#1a9988",
                                "I-76" = "#eb5600",
                                "I-95" = "#eba075",
                                "I-676" = "#7fe3d6")) +
  scale_fill_manual(values = c("US-1" = "#1a9988",
                               "I-76" = "#eb5600",
                               "I-95" = "#eba075",
                               "I-676" = "#7fe3d6")) +
  labs(title = "Property Price (log price per TA; under $1M) as a function of Distance to Closest Highway",
       subtitle = "Colored by Closest Highway; Properties within 500m of highways",
       fill = "Closest Highway",
       color = "Closest Highway") +
  theme_minimal()
```



# Relationship between Housing Price and other variables

First, I filtered the property data in the following ways.

-   Keep properties whose sale_date is between now and Jan. 1st, 2005

-   Remove properties whose sale price is < \$10,000 and > \$25M

-   Remove properties whose total area and total livable area are under 100 sq. ft.

Then, I calculated each property's distance to the four highways of interest as well as amenities. The amenities that we've looked into so far are parks, libraries, transit stops, schools, City Hall (proxy for the central business district), hospitals, museums, and historic sites. We also included distance to prisons as a disamenity.

We still need to look at sale price's relationship with restaurants and grocery stores (I just need to figure out how to use the google API to do this).

```{r cleaning_properties}
# takes about 7 min to run
# cleaned_properties <- philly_properties %>% 
#   filter(sale_price >= 10000,
#          sale_price <= 25000000,
#          total_area >= 100,
#          total_livable_area >= 100,
#          sale_date >= as.Date("2005-01-01", format = "%Y-%m-%d"),
#          sale_date <= Sys.Date()) %>% 
#   mutate(price_perTLA = sale_price / total_livable_area,
#          price_perTA = sale_price / total_area,
#          log_price = log(sale_price),
#          log_price_perTLA = log(price_perTLA),
#          log_price_perTA = log(price_perTA),
#          norm_log_price = scale(log_price),
#          norm_log_price_perTLA = scale(log_price_perTLA),
#          norm_log_price_perTA = scale(log_price_perTA),
#          category_easy = case_when(category_code_description %in% c("APARTMENTS  > 4 UNITS",
#                                                                     "MIXED USE",
#                                                                     "MULTI FAMILY",
#                                                                     "SINGLE FAMILY") ~ "Residential or Mixed",
#                                    grepl("INDUS", category_code_description) ~ "Industrial",
#                                    grepl("HOTEL", category_code_description) ~ "Hotel",
#                                    grepl("OFFICES", category_code_description) ~ "Offices",
#                                    category_code_description %in% c("COMMERCIAL",
#                                                                     "GARAGE - COMMERCIAL",
#                                                                     "RETAIL") ~ "Commercial",
#                                    grepl("VACANT", category_code_description) ~ "Vacant",
#                                    .default = "Other"),
#          dist_to_US001 = calculate_nearest_distance(geometry,
#                                      philly_mainhighways %>%
#                                        filter(ST_RT_NO == "0001")),
#          dist_to_I076 = calculate_nearest_distance(geometry,
#                                     philly_mainhighways %>%
#                                       filter(ST_RT_NO == "0076")),
#          dist_to_I095 = calculate_nearest_distance(geometry,
#                                     philly_mainhighways %>%
#                                       filter(ST_RT_NO == "0095")),
#          dist_to_I676 = calculate_nearest_distance(geometry,
#                                     philly_mainhighways %>%
#                                       filter(ST_RT_NO == "0676")),
#          dist_park = calculate_nearest_distance(geometry,
#                                                 philly_parks),
#          dist_lib = calculate_nearest_distance(geometry,
#                                                libraries),
#          dist_transit = calculate_nearest_distance(geometry,
#                                                    transit),
#          dist_school = calculate_nearest_distance(geometry,
#                                                   school),
#          dist_cityhall = calculate_nearest_distance(geometry,
#                                                     city_hall),
#          dist_hospital = calculate_nearest_distance(geometry,
#                                                     hospitals),
#          dist_museum = calculate_nearest_distance(geometry,
#                                                   museums),
#          dist_prison = calculate_nearest_distance(geometry,
#                                                   prisons),
#          dist_historic = calculate_nearest_distance(geometry,
#                                                     historic),
#          dist_to_US001m = dist_to_US001 * 0.3048,
#          dist_to_I076m = dist_to_I076 * 0.3048,
#          dist_to_I095m = dist_to_I095 * 0.3048,
#          dist_to_I676m = dist_to_I676 * 0.3048,
#          dist_parkm = dist_park * 0.3048,
#          dist_libm = dist_lib * 0.3048,
#          dist_transitm = dist_transit * 0.3048,
#          dist_schoolm = dist_school * 0.3048,
#          dist_cityhallm = dist_cityhall * 0.3048,
#          dist_hospitalm = dist_hospital * 0.3048,
#          dist_museumm = dist_museum * 0.3048,
#          dist_prisonm = dist_prison * 0.3048,
#          dist_historicm = dist_historic * 0.3048)


# find closest highway
# clean_closest <- cleaned_properties %>%
#   select(matches("objectid|dist_to_US|dist_to_I")) %>%
#   pivot_longer(cols = 2:5,
#                names_to = "highway",
#                values_to = "distance") %>%
#   group_by(objectid) %>%
#   summarise(dist_to_closest_highway = min(distance, na.rm = T)) %>%
#   ungroup()  %>% 
#   mutate(dist_to_closest_highwaym = dist_to_closest_highway * 0.3048) 

# join closest highway
# cleaned_props <- left_join(cleaned_properties,
#                            clean_closest %>% st_drop_geometry(),
#                            by = "objectid") %>%
#   mutate(closest_highway = case_when(
#     dist_to_closest_highway == dist_to_US001 ~ "US-1",
#     dist_to_closest_highway == dist_to_I076 ~ "I-76",
#     dist_to_closest_highway == dist_to_I095 ~ "I-95",
#     dist_to_closest_highway == dist_to_I676 ~ "I-676",
#     .default = NA),
#     closest_highway = factor(closest_highway, levels = c("US-1", "I-76", "I-676", "I-95")),
#     highway_type = case_when(closest_highway %in% c("US-1","I-676") ~ "through neighborhood",
#                              closest_highway %in% c("I-95","I-76") ~ "along waterway",
#                              .default = NA)) %>% 
#   filter(!is.na(closest_highway))

# save for easier access later
# st_write(cleaned_props,
#          "/Users/akiradisandro/Documents/MUSA/Spring25/MUSA8010_Practicum/Chinatown/data/philly_cleaned_properties_with_amenities.geojson")

# read in pre-saved data
cleaned_props <- st_read("/Users/akiradisandro/Documents/MUSA/Spring25/MUSA8010_Practicum/Chinatown/data/philly_cleaned_properties_with_amenities.geojson")

```



```{r cleaned_prop_plots}
# plot of log_price_perTA
ggplot() +
  geom_sf(data = philly_bounds, fill = "#c9b887") +
  geom_sf(data = hydro, fill = "#7d97c7", color = "transparent") +
  geom_sf(data = philly_parks, fill = "#6b9169", color = "transparent") +
    geom_sf(data = cleaned_props, aes(color = log_price_perTA), size = 0.2) +
  scale_color_gradient(high = "#eb5600",
                       low = "#f7cfb7") +
  labs(title = "Property Price (log sale price per total area)",
       color = "Log Price") +
  theme_void()


```

I'm going to go through the following columns of interest (amenity predictors) and make a map + scatterplot and get the Pearson correlation coefficient between the amenity/disamenity + sale price (log_price + log_price_perTA).

-   dist_to_closest_highwaym

-   dist_parkm

-   dist_libm

-   dist_transitm

-   dist_schoolm

-   dist_cityhallm

-   dist_hospitalm

-   dist_museumm

-   dist_prisonm

-   dist_historicm

```{r keep_vars_of_interest}
# only keeping two outcomes (log_price + log_price_perTA) and all amenities
vars_of_int <- cleaned_props %>% 
  select(log_price, log_price_perTA, 
         dist_to_closest_highwaym, dist_parkm, dist_libm, dist_transitm, dist_schoolm, dist_cityhallm, dist_hospitalm, dist_museumm, dist_prisonm, dist_historicm)
```

```{r corplot_amenities}
# cor matrix for corplot
cors_logprice <- cor(vars_of_int %>% 
                       select(-log_price_perTA) %>% 
                       rename(dist_2clssthghwy = dist_to_closest_highwaym) %>% 
                       st_drop_geometry())

cors_logpriceperTA <- cor(vars_of_int %>% 
                            select(-log_price) %>% 
                            rename(dist_2clssthghwy = dist_to_closest_highwaym) %>% 
                            st_drop_geometry())

# significance test 
sigtest_logprice <- cor.mtest(vars_of_int %>% 
                                select(-log_price_perTA) %>% 
                                rename(dist_2clssthghwy = dist_to_closest_highwaym) %>% 
                                st_drop_geometry())

sigtest_logpriceperTA <- cor.mtest(vars_of_int %>% 
                                     select(-log_price) %>% 
                                     rename(dist_2clssthghwy = dist_to_closest_highwaym) %>% 
                                     st_drop_geometry())

# pretty correlation matrix of all continuous variables
corrplot(cors_logprice, 
         title = "Correlation Matrix for Log Price",
         p.mat = sigtest_logprice$p, 
         method = 'circle', 
         type = 'lower', 
         insig='blank',
         addCoef.col ='black', 
         number.cex = 0.8, 
         diag=FALSE)

corrplot(cors_logpriceperTA, 
         title = "Correlation Matrix for Log Price",
         p.mat = sigtest_logpriceperTA$p, 
         method = 'circle', 
         type = 'lower', 
         insig='blank',
         addCoef.col ='black', 
         number.cex = 0.8, 
         diag=FALSE)


```

From the correlation matrix, we can see that distance to City Hall, distance to Transit, and distance to museums are all highly collinear with correlation coefficients ranging from .78 (transit and museum) to .98 (museum and city hall). Interestingly, distance to city hall and distance to museum are the variables with the correlation coefficient of highest magnitude with log_price_perTA (at -.38 and -.34 respectively), but distance to transit is not as correlated to the outcome variable (-.18).

When considering all properties in Philadelphia (after cleaning this data a bit), we see that distance to closest highway now has a negative relationship with property sale price (-0.18).


I'm moving forward with log_price_perTA as outcome for now because the size of the property explains a lot of the variance in property sale price already and right now we are focused on external predictors.


### Distance to Parks

```{r dist_park}
dat <- vars_of_int %>% 
  select(log_price_perTA, dist_parkm)

scat_title <- "Log Price (per total area) as a function of Distance to nearest Park"

cor_toplot <- round(cor(dat$dist_parkm, dat$log_price_perTA), 3)

# scatter
ggplot(dat,
       aes(x = dist_parkm,
           y = log_price_perTA)) +
  geom_point(size = 0.2, alpha = 0.3, color = "#1a9988") +
  geom_smooth(method = "lm", color = "#eb5600", fill = "#eba075") +
  geom_text(aes(label = cor_toplot,
                x = 2000,
                y = 10,
                color = "#eb5600"),
            show.legend = F) +
  labs(title = scat_title,
       x = "Distance to nearest Park (m)",
       y = "Log Price (per total area)") +
  theme_minimal()

# map
ggplot() +
  geom_sf(data = philly_bounds, fill = "#c9b887") +
  geom_sf(data = hydro, fill = "#7d97c7", color = "transparent") +
  geom_sf(data = dat, size = 0.2,
          aes(color = dist_parkm)) +
  scale_color_gradient(high = "#f7cfb7",
                       low = "#eb5600") +
  geom_sf(data = philly_parks, fill = "#6b9169", color = "transparent", alpha = 0.75) +
  labs(title = "Distance to nearest park",
       color = "Dist. to park") +
  theme_void()

```
Looking at distance to park without the extreme outlier that is almost 2500m from the nearest park
```{r dist_park_no_outlier}
dat <- vars_of_int %>% 
  select(log_price_perTA, dist_parkm) %>% 
  filter(dist_parkm < 2000)

scat_title <- "Log Price (per total area) as a function of Distance to nearest Park"

cor_toplot <- round(cor(dat$dist_parkm, dat$log_price_perTA), 3)

# scatter
ggplot(dat,
       aes(x = dist_parkm,
           y = log_price_perTA)) +
  geom_point(size = 0.2, alpha = 0.3, color = "#1a9988") +
  geom_smooth(method = "lm", color = "#eb5600", fill = "#eba075") +
  geom_text(aes(label = cor_toplot,
                x = 1500,
                y = 10,
                color = "#eb5600"),
            show.legend = F) +
  labs(title = scat_title,
       subtitle = "Removed outlier",
       x = "Distance to nearest Park (m)",
       y = "Log Price (per total area)") +
  theme_minimal()


```
Removing the outlier had no effect on the correlation.

### Distance to Libraries

```{r dist_lib}
dat <- vars_of_int %>% 
  select(log_price_perTA, dist_libm)

scat_title <- "Log Price (per total area) as a function of Distance to nearest Library"

cor_toplot <- round(cor(dat$dist_libm, dat$log_price_perTA), 3)

# scatter
ggplot(dat,
       aes(x = dist_libm,
           y = log_price_perTA)) +
  geom_point(size = 0.2, alpha = 0.3, color = "#1a9988") +
  geom_smooth(method = "lm", color = "#eb5600", fill = "#eba075") +
  geom_text(aes(label = cor_toplot,
                x = 3000,
                y = 10,
                color = "#eb5600"),
            show.legend = F) +
  labs(title = scat_title,
       x = "Distance to nearest Library (m)",
       y = "Log Price (per total area)") +
  theme_minimal()

# map
ggplot() +
  geom_sf(data = philly_bounds, fill = "#c9b887") +
  geom_sf(data = hydro, fill = "#7d97c7", color = "transparent") +
  geom_sf(data = philly_parks, fill = "#6b9169", color = "transparent") +
  geom_sf(data = dat, size = 0.2,
          aes(color = dist_libm)) +
  geom_sf(data = libraries, color = "#faef73") +
  scale_color_gradient(high = "#f7cfb7",
                       low = "#eb5600") +
  labs(title = "Distance to nearest Library",
       color = "Dist. to Lib.") +
  theme_void()

```

### Distance to Transit

```{r dist_transit}
dat <- vars_of_int %>% 
  select(log_price_perTA, dist_transitm)

scat_title <- "Log Price (per total area) as a function of Distance to nearest Transit Stop"

cor_toplot <- round(cor(dat$dist_transitm, dat$log_price_perTA), 3)

# scatter
ggplot(dat,
       aes(x = dist_transitm,
           y = log_price_perTA)) +
  geom_point(size = 0.2, alpha = 0.3, color = "#1a9988") +
  geom_smooth(method = "lm", color = "#eb5600", fill = "#eba075") +
  geom_text(aes(label = cor_toplot,
                x = 10000,
                y = 10,
                color = "#eb5600"),
            show.legend = F) +
  labs(title = scat_title,
       x = "Distance to nearest Transit Stop (m)",
       y = "Log Price (per total area)") +
  theme_minimal()

# map
ggplot() +
  geom_sf(data = philly_bounds, fill = "#c9b887") +
  geom_sf(data = hydro, fill = "#7d97c7", color = "transparent") +
  geom_sf(data = philly_parks, fill = "#6b9169", color = "transparent") +
  geom_sf(data = dat, size = 0.2,
          aes(color = dist_transitm)) +
  geom_sf(data = transit, color = "#faef73") +
  scale_color_gradient(high = "#f7cfb7",
                       low = "#eb5600") +
  labs(title = "Distance to nearest Transit stop",
       color = "Dist. to station") +
  theme_void()

```

### Distance to Schools

```{r dist_school}
dat <- vars_of_int %>% 
  select(log_price_perTA, dist_schoolm)

scat_title <- "Log Price (per total area) as a function of Distance to nearest School"

cor_toplot <- round(cor(dat$dist_schoolm, dat$log_price_perTA), 3)

# scatter
ggplot(dat,
       aes(x = dist_schoolm,
           y = log_price_perTA)) +
  geom_point(size = 0.2, alpha = 0.3, color = "#1a9988") +
  geom_smooth(method = "lm", color = "#eb5600", fill = "#eba075") +
  geom_text(aes(label = cor_toplot,
                x = 2000,
                y = 10,
                color = "#eb5600"),
            show.legend = F) +
  labs(title = scat_title,
       x = "Distance to nearest School (m)",
       y = "Log Price (per total area)") +
  theme_minimal()

# map
ggplot() +
  geom_sf(data = philly_bounds, fill = "#c9b887") +
  geom_sf(data = hydro, fill = "#7d97c7", color = "transparent") +
  geom_sf(data = philly_parks, fill = "#6b9169", color = "transparent") +
  geom_sf(data = dat, size = 0.2,
          aes(color = dist_schoolm)) +
  geom_sf(data = school, color = "#faef73") +
  scale_color_gradient(high = "#f7cfb7",
                       low = "#eb5600") +
  labs(title = "Distance to nearest School",
       color = "Dist. to School") +
  theme_void()

```

### Distance to City Hall

```{r dist_cityhall}
dat <- vars_of_int %>% 
  select(log_price_perTA, dist_cityhallm)

scat_title <- "Log Price (per total area) as a function of Distance to City Hall"

cor_toplot <- round(cor(dat$dist_cityhallm, dat$log_price_perTA), 3)

# scatter
ggplot(dat,
       aes(x = dist_cityhallm,
           y = log_price_perTA)) +
  geom_point(size = 0.2, alpha = 0.3, color = "#1a9988") +
  geom_smooth(method = "lm", color = "#eb5600", fill = "#eba075") +
  geom_text(aes(label = cor_toplot,
                x = 20000,
                y = 10,
                color = "#eb5600"),
            show.legend = F) +
  labs(title = scat_title,
       x = "Distance to City Hall (m)",
       y = "Log Price (per total area)") +
  theme_minimal()

# map
ggplot() +
  geom_sf(data = philly_bounds, fill = "#c9b887") +
  geom_sf(data = hydro, fill = "#7d97c7", color = "transparent") +
  geom_sf(data = philly_parks, fill = "#6b9169", color = "transparent") +
  geom_sf(data = dat, size = 0.2,
          aes(color = dist_cityhallm)) +
  geom_sf(data = city_hall, color = "#faef73") +
  scale_color_gradient(high = "#f7cfb7",
                       low = "#eb5600") +
  labs(title = "Distance to City Hall",
       color = "Dist. to CH") +
  theme_void()

```

### Distance to Hospitals

```{r dist_hospital}
dat <- vars_of_int %>% 
  select(log_price_perTA, dist_hospitalm)

scat_title <- "Log Price (per total area) as a function of Distance to nearest Hospital"

cor_toplot <- round(cor(dat$dist_hospitalm, dat$log_price_perTA), 3)

# scatter
ggplot(dat,
       aes(x = dist_hospitalm,
           y = log_price_perTA)) +
  geom_point(size = 0.2, alpha = 0.3, color = "#1a9988") +
  geom_smooth(method = "lm", color = "#eb5600", fill = "#eba075") +
  geom_text(aes(label = cor_toplot,
                x = 5000,
                y = 10,
                color = "#eb5600"),
            show.legend = F) +
  labs(title = scat_title,
       x = "Distance to nearest Hospital (m)",
       y = "Log Price (per total area)") +
  theme_minimal()

# map
ggplot() +
  geom_sf(data = philly_bounds, fill = "#c9b887") +
  geom_sf(data = hydro, fill = "#7d97c7", color = "transparent") +
  geom_sf(data = philly_parks, fill = "#6b9169", color = "transparent") +
  geom_sf(data = dat, size = 0.2,
          aes(color = dist_hospitalm)) +
  geom_sf(data = hospitals, color = "#faef73") +
  scale_color_gradient(high = "#f7cfb7",
                       low = "#eb5600") +
  labs(title = "Distance to nearest Hospital",
       color = "Dist. to Hospital") +
  theme_void()

```

### Distance to Museums

```{r dist_museum}
dat <- vars_of_int %>% 
  select(log_price_perTA, dist_museumm)

scat_title <- "Log Price (per total area) as a function of Distance to nearest Museum"

cor_toplot <- round(cor(dat$dist_museumm, dat$log_price_perTA), 3)

# scatter
ggplot(dat,
       aes(x = dist_museumm,
           y = log_price_perTA)) +
  geom_point(size = 0.2, alpha = 0.3, color = "#1a9988") +
  geom_smooth(method = "lm", color = "#eb5600", fill = "#eba075") +
  geom_text(aes(label = cor_toplot,
                x = 15000,
                y = 10,
                color = "#eb5600"),
            show.legend = F) +
  labs(title = scat_title,
       x = "Distance to nearest Museum (m)",
       y = "Log Price (per total area)") +
  theme_minimal()

# map
ggplot() +
  geom_sf(data = philly_bounds, fill = "#c9b887") +
  geom_sf(data = hydro, fill = "#7d97c7", color = "transparent") +
  geom_sf(data = philly_parks, fill = "#6b9169", color = "transparent") +
  geom_sf(data = dat, size = 0.2,
          aes(color = dist_museumm)) +
  geom_sf(data = museums, color = "#faef73") +
  scale_color_gradient(high = "#f7cfb7",
                       low = "#eb5600") +
  labs(title = "Distance to nearest Museum",
       color = "Dist. to Museum") +
  theme_void()

```

### Distance to Prisons

```{r dist_prison}
dat <- vars_of_int %>% 
  select(log_price_perTA, dist_prisonm)

scat_title <- "Log Price (per total area) as a function of Distance to nearest Prison"

cor_toplot <- round(cor(dat$dist_prisonm, dat$log_price_perTA), 3)

# scatter
ggplot(dat,
       aes(x = dist_prisonm,
           y = log_price_perTA)) +
  geom_point(size = 0.2, alpha = 0.3, color = "#1a9988") +
  geom_smooth(method = "lm", color = "#eb5600", fill = "#eba075") +
  geom_text(aes(label = cor_toplot,
                x = 9000,
                y = 10,
                color = "#eb5600"),
            show.legend = F) +
  labs(title = scat_title,
       x = "Distance to nearest Prison (m)",
       y = "Log Price (per total area)") +
  theme_minimal()

# map
ggplot() +
  geom_sf(data = philly_bounds, fill = "#c9b887") +
  geom_sf(data = hydro, fill = "#7d97c7", color = "transparent") +
  geom_sf(data = philly_parks, fill = "#6b9169", color = "transparent") +
  geom_sf(data = dat, size = 0.2,
          aes(color = dist_prisonm)) +
  geom_sf(data = prisons, color = "#faef73") +
  scale_color_gradient(high = "#f7cfb7",
                       low = "#eb5600") +
  labs(title = "Distance to nearest Prison",
       color = "Dist. to Prison") +
  theme_void()

```

### Distance to Historic Site

```{r dist_hist}
dat <- vars_of_int %>% 
  select(log_price_perTA, dist_historicm)

scat_title <- "Log Price (per total area) as a function of Distance to nearest Historic Site"

cor_toplot <- round(cor(dat$dist_historicm, dat$log_price_perTA), 3)

# scatter
ggplot(dat,
       aes(x = dist_historicm,
           y = log_price_perTA)) +
  geom_point(size = 0.2, alpha = 0.3, color = "#1a9988") +
  geom_smooth(method = "lm", color = "#eb5600", fill = "#eba075") +
  geom_text(aes(label = cor_toplot,
                x = 6000,
                y = 10,
                color = "#eb5600"),
            show.legend = F) +
  labs(title = scat_title,
       x = "Distance to nearest Historic Site (m)",
       y = "Log Price (per total area)") +
  theme_minimal()

# map
ggplot() +
  geom_sf(data = philly_bounds, fill = "#c9b887") +
  geom_sf(data = hydro, fill = "#7d97c7", color = "transparent") +
  geom_sf(data = philly_parks, fill = "#6b9169", color = "transparent") +
  geom_sf(data = dat, size = 0.2,
          aes(color = dist_historicm)) +
  geom_sf(data = historic, color = "#faef73") +
  scale_color_gradient(high = "#f7cfb7",
                       low = "#eb5600") +
  labs(title = "Distance to Historic Site",
       color = "Dist. to Hist.") +
  theme_void()

```


# Relationship between Housing Price and other variables (only keeping properties whose sale date is within the last 5 years)

First, I filtered the property data in the following ways.

-   Keep properties whose sale_date is between now and Jan. 1st, 2020

-   Remove properties whose sale price is < \$10,000 and > \$25M

-   Remove properties whose total area and total livable area are under 100 sq. ft.


```{r cleaning_properties}
# read in pre-saved data
cleaned_props <- st_read("/Users/akiradisandro/Documents/MUSA/Spring25/MUSA8010_Practicum/Chinatown/data/philly_cleaned_properties_with_amenities.geojson") %>% 
  filter(sale_date >= as.Date("2005-01-01", format = "%Y-%m-%d"),)

```



```{r cleaned_prop_plots}
# plot of log_price_perTA
ggplot() +
  geom_sf(data = philly_bounds, fill = "#c9b887") +
  geom_sf(data = hydro, fill = "#7d97c7", color = "transparent") +
  geom_sf(data = philly_parks, fill = "#6b9169", color = "transparent") +
    geom_sf(data = cleaned_props, aes(color = log_price_perTA), size = 0.2) +
  scale_color_gradient(high = "#eb5600",
                       low = "#f7cfb7") +
  labs(title = "Property Price (log sale price per total area)",
       color = "Log Price") +
  theme_void()


```

I'm going to go through the following columns of interest (amenity predictors) and make a map + scatterplot and get the Pearson correlation coefficient between the amenity/disamenity + sale price (log_price + log_price_perTA).

-   dist_to_closest_highwaym

-   dist_parkm

-   dist_libm

-   dist_transitm

-   dist_schoolm

-   dist_cityhallm

-   dist_hospitalm

-   dist_museumm

-   dist_prisonm

-   dist_historicm

```{r keep_vars_of_interest}
# only keeping two outcomes (log_price + log_price_perTA) and all amenities
vars_of_int <- cleaned_props %>% 
  select(log_price, log_price_perTA, 
         dist_to_closest_highwaym, dist_parkm, dist_libm, dist_transitm, dist_schoolm, dist_cityhallm, dist_hospitalm, dist_museumm, dist_prisonm, dist_historicm)
```

```{r corplot_amenities}
# cor matrix for corplot
cors_logprice <- cor(vars_of_int %>% 
                       select(-log_price_perTA) %>% 
                       rename(dist_2clssthghwy = dist_to_closest_highwaym) %>% 
                       st_drop_geometry())

cors_logpriceperTA <- cor(vars_of_int %>% 
                            select(-log_price) %>% 
                            rename(dist_2clssthghwy = dist_to_closest_highwaym) %>% 
                            st_drop_geometry())

# significance test 
sigtest_logprice <- cor.mtest(vars_of_int %>% 
                                select(-log_price_perTA) %>% 
                                rename(dist_2clssthghwy = dist_to_closest_highwaym) %>% 
                                st_drop_geometry())

sigtest_logpriceperTA <- cor.mtest(vars_of_int %>% 
                                     select(-log_price) %>% 
                                     rename(dist_2clssthghwy = dist_to_closest_highwaym) %>% 
                                     st_drop_geometry())

# pretty correlation matrix of all continuous variables
corrplot(cors_logprice, 
         title = "Correlation Matrix for Log Price",
         p.mat = sigtest_logprice$p, 
         method = 'circle', 
         type = 'lower', 
         insig='blank',
         addCoef.col ='black', 
         number.cex = 0.8, 
         diag=FALSE)

corrplot(cors_logpriceperTA, 
         title = "Correlation Matrix for Log Price",
         p.mat = sigtest_logpriceperTA$p, 
         method = 'circle', 
         type = 'lower', 
         insig='blank',
         addCoef.col ='black', 
         number.cex = 0.8, 
         diag=FALSE)


```

From the correlation matrix, we can see that distance to City Hall, distance to Transit, and distance to museums are all highly collinear with correlation coefficients ranging from .78 (transit and museum) to .98 (museum and city hall). Interestingly, distance to city hall and distance to museum are the variables with the correlation coefficient of highest magnitude with log_price_perTA (at -.38 and -.34 respectively), but distance to transit is not as correlated to the outcome variable (-.18).

When considering all properties in Philadelphia (after cleaning this data a bit), we see that distance to closest highway now has a negative relationship with property sale price (-0.18).


I'm moving forward with log_price_perTA as outcome for now because the size of the property explains a lot of the variance in property sale price already and right now we are focused on external predictors.


### Distance to Parks

```{r dist_park}
dat <- vars_of_int %>% 
  select(log_price_perTA, dist_parkm)

scat_title <- "Log Price (per total area) as a function of Distance to nearest Park"

cor_toplot <- round(cor(dat$dist_parkm, dat$log_price_perTA), 3)

# scatter
ggplot(dat,
       aes(x = dist_parkm,
           y = log_price_perTA)) +
  geom_point(size = 0.2, alpha = 0.3, color = "#1a9988") +
  geom_smooth(method = "lm", color = "#eb5600", fill = "#eba075") +
  geom_text(aes(label = cor_toplot,
                x = 2000,
                y = 10,
                color = "#eb5600"),
            show.legend = F) +
  labs(title = scat_title,
       x = "Distance to nearest Park (m)",
       y = "Log Price (per total area)") +
  theme_minimal()

# map
ggplot() +
  geom_sf(data = philly_bounds, fill = "#c9b887") +
  geom_sf(data = hydro, fill = "#7d97c7", color = "transparent") +
  geom_sf(data = dat, size = 0.2,
          aes(color = dist_parkm)) +
  scale_color_gradient(high = "#f7cfb7",
                       low = "#eb5600") +
  geom_sf(data = philly_parks, fill = "#6b9169", color = "transparent", alpha = 0.75) +
  labs(title = "Distance to nearest park",
       color = "Dist. to park") +
  theme_void()

```
Looking at distance to park without the extreme outlier that is almost 2500m from the nearest park
```{r dist_park_no_outlier}
dat <- vars_of_int %>% 
  select(log_price_perTA, dist_parkm) %>% 
  filter(dist_parkm < 2000)

scat_title <- "Log Price (per total area) as a function of Distance to nearest Park"

cor_toplot <- round(cor(dat$dist_parkm, dat$log_price_perTA), 3)

# scatter
ggplot(dat,
       aes(x = dist_parkm,
           y = log_price_perTA)) +
  geom_point(size = 0.2, alpha = 0.3, color = "#1a9988") +
  geom_smooth(method = "lm", color = "#eb5600", fill = "#eba075") +
  geom_text(aes(label = cor_toplot,
                x = 1500,
                y = 10,
                color = "#eb5600"),
            show.legend = F) +
  labs(title = scat_title,
       subtitle = "Removed outlier",
       x = "Distance to nearest Park (m)",
       y = "Log Price (per total area)") +
  theme_minimal()


```
Removing the outlier had no effect on the correlation.

### Distance to Libraries

```{r dist_lib}
dat <- vars_of_int %>% 
  select(log_price_perTA, dist_libm)

scat_title <- "Log Price (per total area) as a function of Distance to nearest Library"

cor_toplot <- round(cor(dat$dist_libm, dat$log_price_perTA), 3)

# scatter
ggplot(dat,
       aes(x = dist_libm,
           y = log_price_perTA)) +
  geom_point(size = 0.2, alpha = 0.3, color = "#1a9988") +
  geom_smooth(method = "lm", color = "#eb5600", fill = "#eba075") +
  geom_text(aes(label = cor_toplot,
                x = 3000,
                y = 10,
                color = "#eb5600"),
            show.legend = F) +
  labs(title = scat_title,
       x = "Distance to nearest Library (m)",
       y = "Log Price (per total area)") +
  theme_minimal()

# map
ggplot() +
  geom_sf(data = philly_bounds, fill = "#c9b887") +
  geom_sf(data = hydro, fill = "#7d97c7", color = "transparent") +
  geom_sf(data = philly_parks, fill = "#6b9169", color = "transparent") +
  geom_sf(data = dat, size = 0.2,
          aes(color = dist_libm)) +
  geom_sf(data = libraries, color = "#faef73") +
  scale_color_gradient(high = "#f7cfb7",
                       low = "#eb5600") +
  labs(title = "Distance to nearest Library",
       color = "Dist. to Lib.") +
  theme_void()

```

### Distance to Transit

```{r dist_transit}
dat <- vars_of_int %>% 
  select(log_price_perTA, dist_transitm)

scat_title <- "Log Price (per total area) as a function of Distance to nearest Transit Stop"

cor_toplot <- round(cor(dat$dist_transitm, dat$log_price_perTA), 3)

# scatter
ggplot(dat,
       aes(x = dist_transitm,
           y = log_price_perTA)) +
  geom_point(size = 0.2, alpha = 0.3, color = "#1a9988") +
  geom_smooth(method = "lm", color = "#eb5600", fill = "#eba075") +
  geom_text(aes(label = cor_toplot,
                x = 10000,
                y = 10,
                color = "#eb5600"),
            show.legend = F) +
  labs(title = scat_title,
       x = "Distance to nearest Transit Stop (m)",
       y = "Log Price (per total area)") +
  theme_minimal()

# map
ggplot() +
  geom_sf(data = philly_bounds, fill = "#c9b887") +
  geom_sf(data = hydro, fill = "#7d97c7", color = "transparent") +
  geom_sf(data = philly_parks, fill = "#6b9169", color = "transparent") +
  geom_sf(data = dat, size = 0.2,
          aes(color = dist_transitm)) +
  geom_sf(data = transit, color = "#faef73") +
  scale_color_gradient(high = "#f7cfb7",
                       low = "#eb5600") +
  labs(title = "Distance to nearest Transit stop",
       color = "Dist. to station") +
  theme_void()

```

### Distance to Schools

```{r dist_school}
dat <- vars_of_int %>% 
  select(log_price_perTA, dist_schoolm)

scat_title <- "Log Price (per total area) as a function of Distance to nearest School"

cor_toplot <- round(cor(dat$dist_schoolm, dat$log_price_perTA), 3)

# scatter
ggplot(dat,
       aes(x = dist_schoolm,
           y = log_price_perTA)) +
  geom_point(size = 0.2, alpha = 0.3, color = "#1a9988") +
  geom_smooth(method = "lm", color = "#eb5600", fill = "#eba075") +
  geom_text(aes(label = cor_toplot,
                x = 2000,
                y = 10,
                color = "#eb5600"),
            show.legend = F) +
  labs(title = scat_title,
       x = "Distance to nearest School (m)",
       y = "Log Price (per total area)") +
  theme_minimal()

# map
ggplot() +
  geom_sf(data = philly_bounds, fill = "#c9b887") +
  geom_sf(data = hydro, fill = "#7d97c7", color = "transparent") +
  geom_sf(data = philly_parks, fill = "#6b9169", color = "transparent") +
  geom_sf(data = dat, size = 0.2,
          aes(color = dist_schoolm)) +
  geom_sf(data = school, color = "#faef73") +
  scale_color_gradient(high = "#f7cfb7",
                       low = "#eb5600") +
  labs(title = "Distance to nearest School",
       color = "Dist. to School") +
  theme_void()

```

### Distance to City Hall

```{r dist_cityhall}
dat <- vars_of_int %>% 
  select(log_price_perTA, dist_cityhallm)

scat_title <- "Log Price (per total area) as a function of Distance to City Hall"

cor_toplot <- round(cor(dat$dist_cityhallm, dat$log_price_perTA), 3)

# scatter
ggplot(dat,
       aes(x = dist_cityhallm,
           y = log_price_perTA)) +
  geom_point(size = 0.2, alpha = 0.3, color = "#1a9988") +
  geom_smooth(method = "lm", color = "#eb5600", fill = "#eba075") +
  geom_text(aes(label = cor_toplot,
                x = 20000,
                y = 10,
                color = "#eb5600"),
            show.legend = F) +
  labs(title = scat_title,
       x = "Distance to City Hall (m)",
       y = "Log Price (per total area)") +
  theme_minimal()

# map
ggplot() +
  geom_sf(data = philly_bounds, fill = "#c9b887") +
  geom_sf(data = hydro, fill = "#7d97c7", color = "transparent") +
  geom_sf(data = philly_parks, fill = "#6b9169", color = "transparent") +
  geom_sf(data = dat, size = 0.2,
          aes(color = dist_cityhallm)) +
  geom_sf(data = city_hall, color = "#faef73") +
  scale_color_gradient(high = "#f7cfb7",
                       low = "#eb5600") +
  labs(title = "Distance to City Hall",
       color = "Dist. to CH") +
  theme_void()

```

### Distance to Hospitals

```{r dist_hospital}
dat <- vars_of_int %>% 
  select(log_price_perTA, dist_hospitalm)

scat_title <- "Log Price (per total area) as a function of Distance to nearest Hospital"

cor_toplot <- round(cor(dat$dist_hospitalm, dat$log_price_perTA), 3)

# scatter
ggplot(dat,
       aes(x = dist_hospitalm,
           y = log_price_perTA)) +
  geom_point(size = 0.2, alpha = 0.3, color = "#1a9988") +
  geom_smooth(method = "lm", color = "#eb5600", fill = "#eba075") +
  geom_text(aes(label = cor_toplot,
                x = 5000,
                y = 10,
                color = "#eb5600"),
            show.legend = F) +
  labs(title = scat_title,
       x = "Distance to nearest Hospital (m)",
       y = "Log Price (per total area)") +
  theme_minimal()

# map
ggplot() +
  geom_sf(data = philly_bounds, fill = "#c9b887") +
  geom_sf(data = hydro, fill = "#7d97c7", color = "transparent") +
  geom_sf(data = philly_parks, fill = "#6b9169", color = "transparent") +
  geom_sf(data = dat, size = 0.2,
          aes(color = dist_hospitalm)) +
  geom_sf(data = hospitals, color = "#faef73") +
  scale_color_gradient(high = "#f7cfb7",
                       low = "#eb5600") +
  labs(title = "Distance to nearest Hospital",
       color = "Dist. to Hospital") +
  theme_void()

```

### Distance to Museums

```{r dist_museum}
dat <- vars_of_int %>% 
  select(log_price_perTA, dist_museumm)

scat_title <- "Log Price (per total area) as a function of Distance to nearest Museum"

cor_toplot <- round(cor(dat$dist_museumm, dat$log_price_perTA), 3)

# scatter
ggplot(dat,
       aes(x = dist_museumm,
           y = log_price_perTA)) +
  geom_point(size = 0.2, alpha = 0.3, color = "#1a9988") +
  geom_smooth(method = "lm", color = "#eb5600", fill = "#eba075") +
  geom_text(aes(label = cor_toplot,
                x = 15000,
                y = 10,
                color = "#eb5600"),
            show.legend = F) +
  labs(title = scat_title,
       x = "Distance to nearest Museum (m)",
       y = "Log Price (per total area)") +
  theme_minimal()

# map
ggplot() +
  geom_sf(data = philly_bounds, fill = "#c9b887") +
  geom_sf(data = hydro, fill = "#7d97c7", color = "transparent") +
  geom_sf(data = philly_parks, fill = "#6b9169", color = "transparent") +
  geom_sf(data = dat, size = 0.2,
          aes(color = dist_museumm)) +
  geom_sf(data = museums, color = "#faef73") +
  scale_color_gradient(high = "#f7cfb7",
                       low = "#eb5600") +
  labs(title = "Distance to nearest Museum",
       color = "Dist. to Museum") +
  theme_void()

```

### Distance to Prisons

```{r dist_prison}
dat <- vars_of_int %>% 
  select(log_price_perTA, dist_prisonm)

scat_title <- "Log Price (per total area) as a function of Distance to nearest Prison"

cor_toplot <- round(cor(dat$dist_prisonm, dat$log_price_perTA), 3)

# scatter
ggplot(dat,
       aes(x = dist_prisonm,
           y = log_price_perTA)) +
  geom_point(size = 0.2, alpha = 0.3, color = "#1a9988") +
  geom_smooth(method = "lm", color = "#eb5600", fill = "#eba075") +
  geom_text(aes(label = cor_toplot,
                x = 9000,
                y = 10,
                color = "#eb5600"),
            show.legend = F) +
  labs(title = scat_title,
       x = "Distance to nearest Prison (m)",
       y = "Log Price (per total area)") +
  theme_minimal()

# map
ggplot() +
  geom_sf(data = philly_bounds, fill = "#c9b887") +
  geom_sf(data = hydro, fill = "#7d97c7", color = "transparent") +
  geom_sf(data = philly_parks, fill = "#6b9169", color = "transparent") +
  geom_sf(data = dat, size = 0.2,
          aes(color = dist_prisonm)) +
  geom_sf(data = prisons, color = "#faef73") +
  scale_color_gradient(high = "#f7cfb7",
                       low = "#eb5600") +
  labs(title = "Distance to nearest Prison",
       color = "Dist. to Prison") +
  theme_void()

```

### Distance to Historic Site

```{r dist_hist}
dat <- vars_of_int %>% 
  select(log_price_perTA, dist_historicm)

scat_title <- "Log Price (per total area) as a function of Distance to nearest Historic Site"

cor_toplot <- round(cor(dat$dist_historicm, dat$log_price_perTA), 3)

# scatter
ggplot(dat,
       aes(x = dist_historicm,
           y = log_price_perTA)) +
  geom_point(size = 0.2, alpha = 0.3, color = "#1a9988") +
  geom_smooth(method = "lm", color = "#eb5600", fill = "#eba075") +
  geom_text(aes(label = cor_toplot,
                x = 6000,
                y = 10,
                color = "#eb5600"),
            show.legend = F) +
  labs(title = scat_title,
       x = "Distance to nearest Historic Site (m)",
       y = "Log Price (per total area)") +
  theme_minimal()

# map
ggplot() +
  geom_sf(data = philly_bounds, fill = "#c9b887") +
  geom_sf(data = hydro, fill = "#7d97c7", color = "transparent") +
  geom_sf(data = philly_parks, fill = "#6b9169", color = "transparent") +
  geom_sf(data = dat, size = 0.2,
          aes(color = dist_historicm)) +
  geom_sf(data = historic, color = "#faef73") +
  scale_color_gradient(high = "#f7cfb7",
                       low = "#eb5600") +
  labs(title = "Distance to Historic Site",
       color = "Dist. to Hist.") +
  theme_void()

```
