---
title: "EDA - Aki"
author: "Luming Xu, Sijia Zhang, Aki Di Sandro, Ray Ma, Yixuan Zhou"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
---

```{r r-setup, results='hide', echo = F, eval = T, warning = F, message = F}
knitr::opts_chunk$set(echo = T, eval = T, warning = F, message = F)

# set working directory
setwd("~/Documents/MUSA/Spring25/MUSA8010_Practicum/Chinatown/")

# load all packages
library(tidyverse)
library(tidycensus)
library(sf)
library(ggplot2)

# read in API
census_api_key("f2855a6037284cb9cbed55e96e6b99be17ee05c6", overwrite = TRUE, install = T)

# mapbox API
# options(rdeck.mapbox_access_token = "pk.eyJ1IjoiY2hpYmlha2kiLCJhIjoiY20xODh2NTNqMTBvaDJqb2ptbjM4ZGViayJ9.un9M1_-S6kI8M0ktqZLz_Q")

```

# Loading all data

```{r load_data}
# study area
# studyarea <- st_read("data/StudyArea.shp") %>% 
studyarea <- st_read("~/Documents/MUSA/Spring25/MUSA8010_Practicum/Chinatown/Dataset/studyarea/StudyArea.shp") %>%
  st_transform('EPSG:2272')

# Philly block groups
# blockgroups <- st_read("data/Philly_blockgroup.shp") %>%
blockgroups <- st_read("~/Documents/MUSA/Spring25/MUSA8010_Practicum/Chinatown/Dataset/Philly_blockgroup/Philly_blockgroup.shp") %>%
  st_transform('EPSG:2272')

# philly bounds
philly_bounds <- st_union(blockgroups)

# philly hydrology (bounded by philly_bounds; source: https://opendataphilly.org/datasets/hydrology/)
hydro <- st_read("https://services.arcgis.com/fLeGjb7u4uXqeF9q/arcgis/rest/services/Hydrographic_Features_Poly/FeatureServer/1/query?outFields=*&where=1%3D1&f=geojson") %>% 
    st_transform(crs = 'EPSG:2272') %>% 
  st_intersection(philly_bounds)

# highway
# stateroads_inphilly <- st_read("data/PaStateRoads2024_03.geojson") %>% 
#   st_transform('EPSG:2272') %>% 
#   st_intersection(philly_bounds)

# write out stateroads_inphilly as its own file since it's big
# st_write(stateroads_inphilly,
#          "data/PhillyStateRoads.shp")

# read in pre-filtered stateroads data
stateroads_inphilly <- st_read("~/Documents/MUSA/Spring25/MUSA8010_Practicum/Chinatown/data/PhillyStateRoads.shp")

philly_mainhighways <- stateroads_inphilly %>%
  filter(TRAF_RT_NO %in% c("I", "US"),
         ST_RT_NO %in% c("0001", "0095", "0076", "0676")) %>%
  dplyr::select(STREET_NAM, ST_RT_NO, TRAF_RT_NO, TRAF_RT__1)

# save and only use the highways of interest
# for now, interested in comparing highways that cut through neighborhoods vs those that don't

# ACS 


# philly property data (https://opendataphilly.org/datasets/philadelphia-properties-and-assessment-history/) <-- that was the original source, but now i'm using the same downloaded file that Luming and Sijia have been working from
philly_properties <- st_read("~/Documents/MUSA/Spring25/MUSA8010_Practicum/Chinatown/data/opa_properties_public.geojson") %>% 
  st_transform(crs = 'EPSG:2272')

# philly park data (https://opendataphilly.org/datasets/ppr-properties/)
philly_parks <- st_read("https://opendata.arcgis.com/datasets/d52445160ab14380a673e5849203eb64_0.geojson") %>%
  st_transform(crs = 'EPSG:2272')

# philly schools
school <-
  st_read("https://opendata.arcgis.com/datasets/d46a7e59e2c246c891fbee778759717e_0.geojson") %>%
  st_transform('EPSG:2272')

# city facilities (https://opendataphilly.org/datasets/city-facilities-master-facilities-database/)
facilities <- st_read("https://opendata.arcgis.com/datasets/b3c133c3b15d4c96bcd4d5cc09f19f4e_0.geojson") %>%
  st_transform('EPSG:2272') %>% 
  filter(STATUS == "A") # remove inactive sites

# libraries (from facilities data)
libraries <- facilities %>% 
  filter(ASSET_GROUP1 == "A8")

# produce (LPSS, HPSS -- indicators for produce; low or high produce supply stores)
retail_produce <- st_read("https://opendata.arcgis.com/datasets/53b8a1c653a74c92b2de23a5d7bf04a0_0.geojson") %>%
  st_transform('EPSG:2272')


# checking out what Luming_property.csv looks like
luming_property <- read.csv("~/Documents/MUSA/Spring25/MUSA8010_Practicum/Chinatown/Dataset/Luming_property.csv")
# looking at which properties were saved in luming_property -- Luming and Sijia's properties files are limited to the case study area
# luming_properties <- philly_properties %>% 
#   filter(objectid %in% luming_property$objectid)

# checking out what property_sijia_eda.geojson looks like
sijia_eda <- st_read("~/Documents/MUSA/Spring25/MUSA8010_Practicum/Chinatown/Dataset/property_sijia_eda.geojson")

```

# Initial plots

```{r plots_initial}
# interstate highways in philly
ggplot() +
    geom_sf(data = philly_bounds) +
    geom_sf(data = stateroads_inphilly %>% 
              filter(TRAF_RT_NO == "I"))

# US highways in philly
ggplot() +
    geom_sf(data = philly_bounds) +
    geom_sf(data = stateroads_inphilly %>% 
              filter(TRAF_RT_NO == "US"))

# PA state highways in philly
ggplot() +
    geom_sf(data = philly_bounds) +
    geom_sf(data = stateroads_inphilly %>% 
              filter(TRAF_RT_NO == "PA"))

# all other roads in philly
ggplot() +
    geom_sf(data = philly_bounds) +
    geom_sf(data = stateroads_inphilly %>% 
              filter(!(TRAF_RT_NO %in% c("I", "US", "PA"))))

# I-95, US-1, I-676, and I-76
ggplot() +
  geom_sf(data = philly_bounds, fill = "#b39624", color = "#57470a") +
  geom_sf(data = hydro, fill = "#96dbe3", color = "transparent") +
  geom_sf(data = philly_mainhighways,
          aes(color = ST_RT_NO)) +
  
  # add custom colors
  scale_color_manual(values = c("0001" = "red",
                                "0076" = "blue",
                                "0095" = "purple",
                                "0676" = "pink")) +
  
  labs(title = "Case Study of Highways in Philly",
       subtitle = "US-1, I-76, I-95, and I-676",
       color = "Highways") +
  theme_void()

# looking at study area within philly bounds
ggplot() +
  geom_sf(data = philly_bounds, fill = "#b39624") + 
  geom_sf(data = hydro, fill = "#96dbe3", color = "transparent") +
  geom_sf(data = studyarea, fill = "coral")

# philly parks
ggplot() +
  geom_sf(data = philly_bounds, fill = "#b39624") + 
  geom_sf(data = hydro, fill = "#96dbe3", color = "transparent") +
  geom_sf(data = philly_parks, fill = "darkgreen", color = "transparent")

# philly schools
ggplot() +
  geom_sf(data = philly_bounds, fill = "#b39624") + 
  geom_sf(data = hydro, fill = "#96dbe3", color = "transparent") +
  geom_sf(data = school, color = "brown4") # point data

# libraries
ggplot() +
  geom_sf(data = philly_bounds, fill = "#b39624") + 
  geom_sf(data = hydro, fill = "#96dbe3", color = "transparent") +
  geom_sf(data = libraries, color = "blue4") # point data

# transit / looking at what this category actually consists of....
# looks like there's point geography for parking lots, some random transit stops, and heli pad station
ggplot() +
  geom_sf(data = philly_bounds, fill = "#b39624") + 
  geom_sf(data = hydro, fill = "#96dbe3", color = "transparent") +
  geom_sf(data = facilities %>% 
            filter(ASSET_GROUP1 == "A17",
                   grepl("Parking", ASSET_SUBT1_DESC)), color = "pink1") # point data

ggplot() +
  geom_sf(data = philly_bounds, fill = "#b39624") + 
  geom_sf(data = hydro, fill = "#96dbe3", color = "transparent") +
  geom_sf(data = sijia_eda, color = "purple3")

# prison facilities
ggplot() +
  geom_sf(data = philly_bounds, fill = "#b39624") + 
  geom_sf(data = hydro, fill = "#96dbe3", color = "transparent") +
  geom_sf(data = facilities %>% 
            filter(ASSET_GROUP2 == "A13.3"), color = "lightgrey") # point data

# all philly properties
ggplot() +
  geom_sf(data = philly_bounds, fill = "#b39624") + 
  geom_sf(data = hydro, fill = "#96dbe3", color = "transparent") +
  geom_sf(data = philly_properties, color = "black")

```

# Case Study of Philadelphia's Highways and Housing Price

For all Philly properties, calculate the distance from highway and see how it relates to price.

```{r calc_dist_func}
# taking Luming's distance calculator function
calculate_nearest_distance <- function(set_points, other_layer) {
  nearest_idx <- st_nearest_feature(set_points, other_layer)
  st_distance(set_points, other_layer[nearest_idx, ], by_element = TRUE) %>% as.numeric()
}

```

The following code chunk took too much to load (too many properties to consider), so I'm going to limit it to the properties within 500m (and possibly look at 1000m and 1500m) from highways to look at the immediate effects of highways on property prices.

```{r highwaybuffer}
# only keep properties that are within 500m (1640.42 ft) of highways of interest
buff500 <- 500 * 3.28084
highways_500buffer <- st_union(st_buffer(philly_mainhighways, buff500))

properties500 <- philly_properties[highways_500buffer,]

# only keep properties that are within 1000m (3280.84 ft) of highways of interest
buff1000 <- 1000 * 3.28084
highways_1000buffer <- st_union(st_buffer(philly_mainhighways, buff1000))

properties1000 <- philly_properties[highways_1000buffer,]

# only keep properties that are within 1500m (4921.26 ft) of highways of interest
buff1500 <- 1500 * 3.28084
highways_1500buffer <- st_union(st_buffer(philly_mainhighways, buff1500))

properties1500 <- philly_properties[highways_1500buffer,]

```

```{r temp_plots}
# visualizing buffer
ggplot() +
  geom_sf(data = philly_bounds, fill = "#b39624", color = "#57470a") +
  geom_sf(data = hydro, fill = "#96dbe3", color = "transparent") +
  geom_sf(data = highways_500buffer,
          color = "transparent", fill = "grey", alpha = 0.4) +
  geom_sf(data = philly_mainhighways,
          aes(color = ST_RT_NO)) +
  
  # add custom colors
  scale_color_manual(values = c("0001" = "red",
                                "0076" = "blue",
                                "0095" = "purple",
                                "0676" = "pink")) +
  
  labs(title = "Case Study of Highways in Philly",
       subtitle = "US-1, I-76, I-95, and I-676",
       color = "Highways",
       fill = "Highways") +
  theme_void()

# visualizing properties to make sure we got the right ones
ggplot() +
  geom_sf(data = philly_bounds, fill = "#b39624", color = "#57470a") +
  geom_sf(data = hydro, fill = "#96dbe3", color = "transparent") +
  geom_sf(data = philly_parks, fill = "darkgreen", color = "transparent") +
  geom_sf(data = highways_500buffer,
          color = "transparent", fill = "grey", alpha = 0.4) +
  geom_sf(data = properties500,
          color = "#4a4a4a", size = .3) +
  geom_sf(data = philly_mainhighways,
          aes(color = ST_RT_NO)) +
  
  # add custom colors
  scale_color_manual(values = c("0001" = "red",
                                "0076" = "blue",
                                "0095" = "purple",
                                "0676" = "pink")) +
  
  labs(title = "Properties within 500m of Highways in Philly",
       subtitle = "US-1, I-76, I-95, and I-676",
       color = "Highways",
       fill = "Highways") +
  theme_void()

ggplot() +
  geom_sf(data = philly_bounds, fill = "#b39624", color = "#57470a") +
  geom_sf(data = hydro, fill = "#96dbe3", color = "transparent") +
  geom_sf(data = philly_parks, fill = "darkgreen", color = "transparent") +
  geom_sf(data = highways_1000buffer,
          color = "transparent", fill = "grey", alpha = 0.4) +
  geom_sf(data = properties1000,
          color = "#4a4a4a", size = .3) +
  geom_sf(data = philly_mainhighways,
          aes(color = ST_RT_NO)) +
  
  # add custom colors
  scale_color_manual(values = c("0001" = "red",
                                "0076" = "blue",
                                "0095" = "purple",
                                "0676" = "pink")) +
  
  labs(title = "Properties within 1000m of Highways in Philly",
       subtitle = "US-1, I-76, I-95, and I-676",
       color = "Highways",
       fill = "Highways") +
  theme_void()

ggplot() +
  geom_sf(data = philly_bounds, fill = "#b39624", color = "#57470a") +
  geom_sf(data = hydro, fill = "#96dbe3", color = "transparent") +
  geom_sf(data = philly_parks, fill = "darkgreen", color = "transparent") +
  geom_sf(data = highways_1500buffer,
          color = "transparent", fill = "grey", alpha = 0.4) +
  geom_sf(data = properties1500,
          color = "#4a4a4a", size = .3) +
  geom_sf(data = philly_mainhighways,
          aes(color = ST_RT_NO)) +
  
  # add custom colors
  scale_color_manual(values = c("0001" = "red",
                                "0076" = "blue",
                                "0095" = "purple",
                                "0676" = "pink")) +
  
  labs(title = "Properties within 1500m of Highways in Philly",
       subtitle = "US-1, I-76, I-95, and I-676",
       color = "Highways",
       fill = "Highways") +
  theme_void()

```

Are there differences between how price of property relates to dist_to_closest_highway depending on type of highway (whether highway cuts through neighborhood or not)?

```{r clean_prop500}
# starting with 500m buffer
# need to check if there are 0 values in sale_price before log transforming
summary(properties500$sale_price)
# Min.   1st Qu.    Median      Mean   3rd Qu.      Max.      NA's 
#   0        10     89900    327917    230000 342000000       840 

# only keep properties whose sale_price >= $10,000
# according to redfin, the most expensive house in philadelphia right now is $25,000,000 -- using that as a cutoff
properties500_clean <- properties500 %>% 
  filter(sale_price >= 10000,
         sale_price <= 25000000,
         total_area >= 100,
         total_livable_area >= 100) %>% 
  mutate(price_perTLA = sale_price / total_livable_area,
         price_perTA = sale_price / total_area,
         log_price = log(sale_price),
         log_price_perTLA = log(price_perTLA),
         log_price_perTA = log(price_perTA),
         norm_log_price = scale(log_price),
         norm_log_price_perTLA = scale(log_price_perTLA),
         norm_log_price_perTA = scale(log_price_perTA))

summary(properties500_clean$sale_price)
#  Min.  1st Qu.   Median     Mean  3rd Qu.     Max. 
# 10000    70000   153000   329289   300000 24973920
summary(properties500_clean$price_perTLA)
# Min.  1st Qu.   Median     Mean  3rd Qu.     Max.     NA's 
# 0.0865  54.8221 117.1875      Inf 231.4815      Inf     1741
summary(properties500_clean$price_perTA)

# correlations between total_area and total_livable_area, and sale_price
# cor(properties500_clean$total_area, properties500_clean$total_livable_area, use = "pairwise.complete.obs") # 0.3501135 -- why is this so low?
# cor(properties500_clean$total_area, properties500_clean$sale_price, use = "pairwise.complete.obs") # 0.2933512
# cor(properties500_clean$total_livable_area, properties500_clean$sale_price, use = "pairwise.complete.obs") # 0.3417108

# histogram of sale_price
ggplot(data = properties500_clean) +
  geom_histogram(aes(x = sale_price),
                 bins = 100)

ggplot(data = properties500_clean) +
  geom_histogram(aes(x = price_perTA),
                 bins = 100)

ggplot(data = properties500_clean) +
  geom_histogram(aes(x = price_perTLA),
                 bins = 100)

```


```{r maps_prop500}
# map of properties within 500 m of highway, colored by sale_price
ggplot() +
  geom_sf(data = philly_bounds, fill = "#c9b887", color = "grey") +
  geom_sf(data = hydro, fill = "#96dbe3", color = "transparent") +
  geom_sf(data = philly_parks, fill = "darkgreen", color = "transparent") +
  geom_sf(data = highways_500buffer,
          color = "transparent", fill = "grey", alpha = 0.4) +
  geom_sf(data = properties500_clean,
          aes(color = sale_price), size = .3) +
  geom_sf(data = philly_mainhighways,
          color = "black") +
  labs(title = "Sale Price of Properties within 500m of Highways in Philly",
       subtitle = "US-1, I-76, I-95, and I-676") +
  theme_void()

# map of properties within 500 m of highway, colored by sale_price
ggplot() +
  geom_sf(data = philly_bounds, fill = "#c9b887", color = "grey") +
  geom_sf(data = hydro, fill = "#96dbe3", color = "transparent") +
  geom_sf(data = philly_parks, fill = "darkgreen", color = "transparent") +
  geom_sf(data = highways_500buffer,
          color = "transparent", fill = "grey", alpha = 0.4) +
  geom_sf(data = properties500_clean,
          aes(color = price_perTA), size = .3) +
  geom_sf(data = philly_mainhighways,
          color = "black") +
  labs(title = "Sale Price (per Total Area) of Properties within 500m of Highways in Philly",
       subtitle = "US-1, I-76, I-95, and I-676") +
  theme_void()

ggplot() +
  geom_sf(data = philly_bounds, fill = "#c9b887", color = "grey") +
  geom_sf(data = hydro, fill = "#96dbe3", color = "transparent") +
  geom_sf(data = philly_parks, fill = "darkgreen", color = "transparent") +
  geom_sf(data = highways_500buffer,
          color = "transparent", fill = "grey", alpha = 0.4) +
  geom_sf(data = properties500_clean,
          aes(color = price_perTLA), size = .3) +
  geom_sf(data = philly_mainhighways,
          color = "black") +
  labs(title = "Sale Price (per Total Livable Area) of Properties within 500m of Highways in Philly",
       subtitle = "US-1, I-76, I-95, and I-676") +
  theme_void()

ggplot() +
  geom_sf(data = philly_bounds, fill = "#c9b887", color = "grey") +
  geom_sf(data = hydro, fill = "#96dbe3", color = "transparent") +
  geom_sf(data = philly_parks, fill = "darkgreen", color = "transparent") +
  geom_sf(data = highways_500buffer,
          color = "transparent", fill = "grey", alpha = 0.4) +
  geom_sf(data = properties500_clean,
          aes(color = log_price), size = .3) +
  geom_sf(data = philly_mainhighways,
          color = "black") +
  labs(title = "Log Sale Price of Properties within 500m of Highways in Philly",
       subtitle = "US-1, I-76, I-95, and I-676") +
  theme_void()

ggplot() +
  geom_sf(data = philly_bounds, fill = "#c9b887", color = "grey") +
  geom_sf(data = hydro, fill = "#96dbe3", color = "transparent") +
  geom_sf(data = philly_parks, fill = "darkgreen", color = "transparent") +
  geom_sf(data = highways_500buffer,
          color = "transparent", fill = "grey", alpha = 0.4) +
  geom_sf(data = properties500_clean,
          aes(color = log_price_perTA), size = .3) +
  geom_sf(data = philly_mainhighways,
          color = "black") +
  labs(title = "Log Sale Price (per Total Area) of Properties within 500m of Highways in Philly",
       subtitle = "US-1, I-76, I-95, and I-676") +
  theme_void()

ggplot() +
  geom_sf(data = philly_bounds, fill = "#c9b887", color = "grey") +
  geom_sf(data = hydro, fill = "#96dbe3", color = "transparent") +
  geom_sf(data = philly_parks, fill = "darkgreen", color = "transparent") +
  geom_sf(data = highways_500buffer,
          color = "transparent", fill = "grey", alpha = 0.4) +
  geom_sf(data = properties500_clean,
          aes(color = norm_log_price_perTA), size = .3) +
  geom_sf(data = philly_mainhighways,
          color = "black") +
  labs(title = "Scaled Log Sale Price (per Total Area) of Properties within 500m of Highways in Philly",
       subtitle = "US-1, I-76, I-95, and I-676") +
  theme_void()

ggplot() +
  geom_sf(data = philly_bounds, fill = "#c9b887", color = "grey") +
  geom_sf(data = hydro, fill = "#96dbe3", color = "transparent") +
  geom_sf(data = philly_parks, fill = "darkgreen", color = "transparent") +
  geom_sf(data = highways_500buffer,
          color = "transparent", fill = "grey", alpha = 0.4) +
  geom_sf(data = properties500_clean,
          aes(color = log_price_perTLA), size = .3) +
  geom_sf(data = philly_mainhighways,
          color = "black") +
  labs(title = "Log Sale Price (per Total Livable Area) of Properties within 500m of Highways in Philly",
       subtitle = "US-1, I-76, I-95, and I-676") +
  theme_void()


```

Calculate each property's distance to highways.

```{r calc_dist2highway}
# calculate each property's distance to highways of interest
# distance is in feet
start_time <- Sys.time()
prop500 <- properties500_clean %>% 
  select(matches("assessment_date|building_code|category_code|census_tract|geographic|zip|market_value|parcel|sale_|taxable|_area|year_built|objectid|geometry|price")) %>% 
  mutate(dist_to_US001 = calculate_nearest_distance(geometry,
                                     philly_mainhighways %>%
                                       filter(ST_RT_NO == "0001")),
         dist_to_I076 = calculate_nearest_distance(geometry,
                                    philly_mainhighways %>%
                                      filter(ST_RT_NO == "0076")),
         dist_to_I095 = calculate_nearest_distance(geometry,
                                    philly_mainhighways %>%
                                      filter(ST_RT_NO == "0095")),
         dist_to_I676 = calculate_nearest_distance(geometry,
                                    philly_mainhighways %>%
                                      filter(ST_RT_NO == "0676"))) # Time difference of 55.20697 secs

# find closest highway
prop500_closest <- prop500 %>%
  select(matches("objectid|dist_to")) %>%
  pivot_longer(cols = 2:5,
               names_to = "highway",
               values_to = "distance") %>%
  group_by(objectid) %>%
  summarise(dist_to_closest_highway = min(distance, na.rm = T)) %>%
  ungroup() # Time difference of 1.736099 mins

# join closest highway
prop500 <- left_join(prop500,
                     prop500_closest %>% st_drop_geometry(),
                     by = "objectid") %>%
  mutate(closest_highway = case_when(
    dist_to_closest_highway == dist_to_US001 ~ "US-1",
    dist_to_closest_highway == dist_to_I076 ~ "I-76",
    dist_to_closest_highway == dist_to_I095 ~ "I-95",
    dist_to_closest_highway == dist_to_I676 ~ "I-676",
    .default = NA))

# save file
# st_write(prop500, "/Users/akiradisandro/Documents/MUSA/Spring25/MUSA8010_Practicum/Chinatown/data/philly_prop500_dist2highway.geojson")

# read in pre-saved file
# prop500_new <- st_read("/Users/akiradisandro/Documents/MUSA/Spring25/MUSA8010_Practicum/Chinatown/Dataset/data/philly_prop500_dist2highway.geojson")

```

Now, look at relationship between distance to highway and price.

```{r price_dist2highway}
# map of properties within 500m of highway colored by which highway they're closest to
ggplot() +
  geom_sf(data = philly_bounds, fill = "#c9b887", color = "grey") +
  geom_sf(data = hydro, fill = "#96dbe3", color = "transparent") +
  geom_sf(data = philly_parks, fill = "darkgreen", color = "transparent") +
  geom_sf(data = highways_500buffer,
          color = "transparent", fill = "grey", alpha = 0.4) +
  geom_sf(data = prop500,
          aes(color = closest_highway), size = 0.3) +
  geom_sf(data = philly_mainhighways,
          color = "black") +
  
  # add custom colors
  scale_color_manual(values = c("US-1" = "red",
                                "I-76" = "blue",
                                "I-95" = "purple",
                                "I-676" = "pink")) +
  
  labs(title = "Properties within 500m of Highways in Philly",
       subtitle = "US-1, I-76, I-95, and I-676") +
  theme_void()

# simple scatter of price and dist2closesthighway
ggplot(data = prop500, 
       aes(x = dist_to_closest_highway, y = sale_price, 
           color = closest_highway, fill = closest_highway)) +
  geom_point(alpha = .2) +
  geom_smooth(method = "lm") +
  scale_color_manual(values = c("US-1" = "red",
                                "I-76" = "blue",
                                "I-95" = "purple",
                                "I-676" = "pink")) +
  scale_fill_manual(values = c("US-1" = "red",
                               "I-76" = "blue",
                               "I-95" = "purple",
                               "I-676" = "pink")) +
  labs(title = "Property Price as a function of Distance to Closest Highway",
       subtitle = "Colored by Closest Highway; Properties within 500m of highways",
       fill = "Closest Highway",
       color = "Closest Highway") +
  theme_minimal()

# simple scatter of price (as log price per total area in sq ft) and dist2closesthighway
ggplot(data = prop500, 
       aes(x = dist_to_closest_highway, y = log_price_perTA, 
           color = closest_highway, fill = closest_highway)) +
  geom_point(alpha = .2) +
  geom_smooth(method = "lm") +
  scale_color_manual(values = c("US-1" = "red",
                                "I-76" = "blue",
                                "I-95" = "purple",
                                "I-676" = "pink")) +
  scale_fill_manual(values = c("US-1" = "red",
                               "I-76" = "blue",
                               "I-95" = "purple",
                               "I-676" = "pink")) +
  labs(title = "Property Price (log price per TA) as a function of Distance to Closest Highway",
       subtitle = "Colored by Closest Highway; Properties within 500m of highways",
       fill = "Closest Highway",
       color = "Closest Highway") +
  theme_minimal()

# simple regression of price (log_price_perTA) and dist2closesthighway
fit1 <- lm(log_price_perTA ~ dist_to_closest_highway, data = prop500)

summary(fit1)

# 

```


# Relationship between Housing Price and other variables

Notably, I'm exploring the relationship between housing price and distance to highways, distance to parks, distance to commercial areas, distance to schools.

For now, I'm using the study area-related data that Luming and Sijia compiled

```{r price_otheramenities}

```

