---
title: "Feature Engineering"
author: "Luming"
date: "`r Sys.Date()`"
output:
    html_document:
      code_folding: hide
---

```{r setup, message = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, cache = FALSE, fig.show = "asis",
    fig.align = "center")
knitr::opts_knit$set(root.dir = "E:/Spring/Practicum/DataAnalysis/Chinatown")

palette5 <- colorRampPalette(c("#1a9988", "#eb5600"))(5)
library(ggplot2)
library(tidyverse)
library(sf)
library(dplyr)
library(tidyr)
library(rlang)  # Needed for `sym()`
library(patchwork)
```

## Set Up

```{r quiet = TRUE, results = 'hide'}
property_EDA <- st_read("Dataset/property_sijia_eda.geojson") %>%
  st_transform('EPSG:2272')

studyarea_north <- st_read("Dataset/studyarea-sub/studyarea_north.shp") %>%
  st_transform('EPSG:2272')

studyarea_south <- st_read("Dataset/studyarea-sub/studyarea_south.shp") %>%
  st_transform('EPSG:2272')

property_highway <- 
  rbind(
    property_EDA %>% st_intersection(studyarea_north["geometry"]) %>%
      mutate(I676 = "north"),
    property_EDA %>% st_intersection(studyarea_south["geometry"]) %>%
      mutate(I676 = "south")
  )
```


```{r}
property_FE_lm <- property_highway %>%
  st_drop_geometry() %>%
  dplyr::select(
    sale_price.x,
    I676,
    distance_to_city_hall,
    distance_to_nearest_transit,
    distance_to_nearest_hospital,
    distance_to_nearest_school,
    distance_to_nearest_park,
    distance_to_nearest_water,
    distance_to_nearest_bikelane,
    distance_to_I676,
    nhoods_name,
    landuse,
    LPSS_PER1000,
    HPSS_PER1000,
    crime_nn1,
    crime_nn2,
    crime_nn3,
    crime_nn4,
    crime_nn5
) %>%
  filter(sale_price.x > 10 & sale_price.x < 1e6)
```

## Continuous Features

```{r quiet = TRUE}
numeric_vars <- property_FE_lm %>% 
    dplyr::select(
      sale_price.x,
      distance_to_city_hall,
      distance_to_nearest_transit,
      distance_to_nearest_hospital,
      distance_to_nearest_school,
      distance_to_nearest_park,
      distance_to_nearest_water,
      distance_to_nearest_bikelane,
      distance_to_I676,
      crime_nn1,
      crime_nn2,
      crime_nn3,
      crime_nn4,
      crime_nn5) %>%  # Select only numeric columns
    colnames()


log_transform_and_plot <- function(df) {
  
  for (var in numeric_vars) {
    original_values <- df[[var]]
    
    # Handle zero values & NA safely
    transformed_values <- ifelse(!is.na(original_values) & original_values == 0, 
                                 log(original_values + 1), 
                                 log(original_values))
    
    # Plot original variable
    hist(original_values, breaks = 50, main = paste("Original:", var), xlab = var, col = "lightblue")
    
    # Plot log-transformed variable
    hist(transformed_values, breaks = 50, main = paste("Log-transformed:", var), xlab = paste("Log(", var, ")"), col = "orange")
    
    par(mfrow = c(1, 2))  # Reset plot layout
    
  }
  
  # Return dataframe with new log-transformed columns
  df_new <- df %>%
    mutate(across(all_of(numeric_vars), 
                  ~ ifelse(!is.na(.) & . == 0, log(. + 1), log(.)), 
                  .names = "log_{.col}"))
  
  invisible(df_new)

}

# Call the function on your dataset
log_transform_and_plot(property_FE_lm)
property_FE_lm_transformed <- log_transform_and_plot(property_FE_lm)

```



```{r}
# Get variable names, excluding specified columns
vars <- property_FE_lm_transformed %>%
  dplyr::select(-sale_price.x, -log_sale_price.x, -I676) %>%
  colnames()

# Identify log-transformed variables
log_vars <- vars[grepl("^log_", vars)]

# Find corresponding original variables by removing "log_" prefix
original_vars <- gsub("^log_", "", log_vars)

# Compute correlation for original and log-transformed variables
correlation.cor <- property_FE_lm_transformed %>%
  dplyr::select(all_of(c(original_vars, log_vars, "log_sale_price.x"))) %>%
  gather(Variable, Value, -log_sale_price.x) %>%
  mutate(Type = ifelse(grepl("^log_", Variable), "Log", "Original")) %>%
  mutate(Variable = gsub("^log_", "", Variable)) %>%
  group_by(Variable) %>%
  summarize(
    Correlation_Original = cor(Value[Type == "Original"], log_sale_price.x[Type == "Original"], use = "complete.obs"),
    Correlation_Log = cor(Value[Type == "Log"], log_sale_price.x[Type == "Log"], use = "complete.obs")
  ) %>%
  ungroup()

correlation.cor

# Loop through variables and create paired plots
for (i in seq_along(original_vars)) {
  
  orig_var <- original_vars[i]  # Original variable name
  log_var <- log_vars[i]        # Corresponding log-transformed variable
  
  # Check if both exist in the dataset
  if (orig_var %in% colnames(property_FE_lm_transformed) & log_var %in% colnames(property_FE_lm_transformed)) {
    
    # Select and reshape data
    df_plot <- property_FE_lm_transformed %>%
      dplyr::select(all_of(c(orig_var, log_var, "log_sale_price.x", "I676"))) %>%
      gather(Variable, Value, -log_sale_price.x, -I676)
    
    # Generate the plot
    p <- ggplot(df_plot, aes(Value, log_sale_price.x, color = factor(I676))) +
      scale_colour_manual(values = c("#eb5600", "#1a9988")) +
      geom_point(size = .5) + 
      geom_smooth(method = "lm", se = FALSE, colour = "black") +
      facet_wrap(~Variable, ncol = 2, scales = "free") +  # Put original and log version side by side
      labs(title = "") +
      theme_minimal() +
      theme(plot.title = element_text(size = 18, face = "bold"),
            strip.text = element_text(size = 10),
            axis.text = element_text(size = 8),
            strip.text.x = element_text(size = 12, face = "italic"),
            strip.background = element_rect(fill = "transparent", color = "transparent"))
    
    # Print the plot
    print(p)
    
    # Optionally save the plot
    # ggsave(filename = paste0("plot_", orig_var, "_log.png"), plot = p, width = 8, height = 4)
  }
}
```

### Selected variables

+ **log_**sale_price.x,
+ distance_to_city_hall,
+ distance_to_nearest_transit,
+ distance_to_nearest_hospital,
+ distance_to_nearest_school,
+ **log_**distance_to_nearest_park,
+ distance_to_nearest_water,
+ **log_**distance_to_nearest_bikelane,
+ distance_to_I676,
+ crime_nn5

```{r}
# stat <- correlation.cor_selected %>% 
#   rowwise() %>%
#   mutate(
#     Mean = ifelse(Field.Name %in% selected_vars, round(mean(stat_philly_data[[Field.Name]], na.rm = TRUE), 3), NA),
#     StDev = ifelse(Field.Name %in% selected_vars, round(sd(stat_philly_data[[Field.Name]], na.rm = TRUE), 3), NA),
#     Min = ifelse(Field.Name %in% selected_vars, round(min(stat_philly_data[[Field.Name]], na.rm = TRUE), 3), NA),
#     Max = ifelse(Field.Name %in% selected_vars, round(max(stat_philly_data[[Field.Name]], na.rm = TRUE), 3), NA)
#   ) %>%
#   ungroup()
```

```{r fig.height=8, fig.width=12}
property_FE_lm_selected <- property_FE_lm_transformed %>%
  dplyr::select(sale_price.x, I676,
                distance_to_city_hall,
                distance_to_nearest_transit,
                distance_to_nearest_hospital,
                distance_to_nearest_school,
                log_distance_to_nearest_park,
                distance_to_nearest_water,
                log_distance_to_nearest_bikelane,
                distance_to_I676,
                crime_nn5)

# plot correlation
correlation.long <- property_FE_lm_selected %>%
  gather(Variable, Value, -sale_price.x, -I676)

correlation.cor_selected <-
  correlation.long %>%
  group_by(Variable) %>%
  summarize(correlation = cor(Value, sale_price.x, use = "complete.obs"))
    
ggplot(correlation.long, aes(Value, sale_price.x, color = factor(I676))) +
  geom_point(size = 0.1) +
  scale_colour_manual(values = c("#eb5600", "#1a9988")) +
  geom_text(data = correlation.cor_selected, aes(label = paste("r =", round(correlation, 2))),
            inherit.aes = FALSE,
            x=-Inf, y=-Inf, vjust = -1, hjust = -.5) +
  geom_smooth(method = "lm", se = FALSE, colour = 'black') +
  facet_wrap(~Variable, ncol = 4, scales = "free") +
  labs(title = "Sales Price as a Function of Continuous Features",
       caption = "Figure ",
       y = "Sales Price ($)",
       x = "Values of Continuous Features") +
  theme_minimal() +
  theme(plot.title = element_text(size = 18, face = "bold"),
        legend.position = "none",
        # panel.border = element_rect(colour = "black", fill=NA, size=1),
        strip.text = element_text(size=10),
        axis.text = element_text(size=8),
        strip.text.x = element_text(size = 12,face="italic"),
        strip.background = element_rect(fill = "transparent", color = "transparent"),)
```


```{r fig.height=8, fig.width=12}
property_FE_lm_selected <- property_FE_lm_transformed %>%
  dplyr::select(log_sale_price.x, I676,
                distance_to_city_hall,
                distance_to_nearest_transit,
                distance_to_nearest_hospital,
                distance_to_nearest_school,
                log_distance_to_nearest_park,
                distance_to_nearest_water,
                log_distance_to_nearest_bikelane,
                distance_to_I676,
                crime_nn5)

# plot correlation
correlation.long <- property_FE_lm_selected %>%
  gather(Variable, Value, -log_sale_price.x, -I676)

correlation.cor_selected <-
  correlation.long %>%
  group_by(Variable) %>%
  summarize(correlation = cor(Value, log_sale_price.x, use = "complete.obs"))
    
ggplot(correlation.long, aes(Value, log_sale_price.x, color = factor(I676))) +
  geom_point(size = 0.1) +
  scale_colour_manual(values = c("#eb5600", "#1a9988")) +
  geom_text(data = correlation.cor_selected, aes(label = paste("r =", round(correlation, 2))),
            inherit.aes = FALSE,
            x=-Inf, y=-Inf, vjust = -1, hjust = -.5) +
  geom_smooth(method = "lm", se = FALSE, colour = 'black') +
  facet_wrap(~Variable, ncol = 4, scales = "free") +
  labs(title = "Log-transformed Sales Price as a Function of Continuous Features",
       caption = "Figure ",
       y = "Log-transformed Sales Price ($)",
       x = "Values of Continuous Features") +
  theme_minimal() +
  theme(plot.title = element_text(size = 18, face = "bold"),
        legend.position = "none",
        # panel.border = element_rect(colour = "black", fill=NA, size=1),
        strip.text = element_text(size=10),
        axis.text = element_text(size=8),
        strip.text.x = element_text(size = 12,face="italic"),
        strip.background = element_rect(fill = "transparent", color = "transparent"),)
```


## Categorical Features

```{r}
factor_vars <- property_FE_lm_transformed %>% 
    dplyr::select(
      nhoods_name,
      landuse,
      LPSS_PER1000,
      HPSS_PER1000) %>%  # Select only numeric columns
    colnames()

property_FE_lm_transformed <- property_FE_lm_transformed %>%
  mutate(LPSS_PER1000 = as.factor(LPSS_PER1000),
             HPSS_PER1000 = as.factor(HPSS_PER1000))

for (var in factor_vars) {
    count_data <- property_FE_lm_transformed %>%
      group_by(!!sym(var)) %>%
      summarise(count = n())
    
    mean_price_data <- property_FE_lm_transformed %>%
      group_by(!!sym(var)) %>%
      summarise(mean_log_price = mean(log_sale_price.x, na.rm = TRUE))
    
    
    plot1 <- ggplot(count_data, aes(x = !!sym(var), y = count)) +
      geom_bar(stat = "identity", fill = "#1a9988", alpha = 0.8, width = 0.4) +
      labs(title = paste("Count of", var)) +
      theme_minimal() +
      theme(
        legend.position = "none", 
        plot.title = element_text(size = 12),
        axis.title.x = element_blank(),
        axis.text.x = element_text(angle = 45, hjust = 1)
      )
    
    plot2 <- ggplot(mean_price_data, aes(x = !!sym(var), y = mean_log_price)) +
      geom_bar(stat = "identity", fill = "#eb5600", alpha = 0.8, width = 0.4) +
      labs(title = paste("Mean Log Price by", var)) +
      theme_minimal() +
      theme(
        legend.position = "none", 
        plot.title = element_text(size = 12),
        axis.title.x = element_blank(),
        axis.text.x = element_text(angle = 45, hjust = 1)
      ) 

    
    print(plot1 | plot2)
    
  }

```

### After grouping

```{r}
property_FE_lm_grouped <- property_FE_lm_transformed %>% 
  mutate(
    nhoods_name = case_when(
      nhoods_name %in% c("CENTER_CITY", "LOGAN_SQUARE") ~ "WEST_PART",
      TRUE ~ nhoods_name
    ),
    landuse = case_when(
      landuse %in% c("Park/Open Space", "Other/Unknown", "Commercial Business/Professional") ~ "Open Space/Business",
      landuse %in% c("Commercial Consumer", "Commercial Mixed Residential",
                     "Industrial", "Transportation", "Culture/Amusement") ~ "Commercial",
      landuse %in% c("Civic/Institution", "Residential High") ~ "Residential High/Civic",
      TRUE ~ landuse
    ),
    LPSS_PER1000 = case_when(
      LPSS_PER1000 %in% c("32.6549929", "44.72605293") ~ "low",
      LPSS_PER1000 %in% c("50.10660981") ~ "mid",
      LPSS_PER1000 %in% c("60.70941337", "131.86813187") ~ "high"
    ),
    HPSS_PER1000 = case_when(
      HPSS_PER1000 %in% c("4.02271652", "5.59701493") ~ "low",
      HPSS_PER1000 %in% c("6.61572866") ~ "mid",
      HPSS_PER1000 %in% c("22.52747253", "7.33287858") ~ "high"
    )
  )
```

```{r}
for (var in factor_vars) {
    count_data <- property_FE_lm_grouped %>%
      group_by(!!sym(var)) %>%
      summarise(count = n())
    
    mean_price_data <- property_FE_lm_grouped %>%
      group_by(!!sym(var)) %>%
      summarise(mean_log_price = mean(log_sale_price.x, na.rm = TRUE))
    
    
    plot1 <- ggplot(count_data, aes(x = !!sym(var), y = count)) +
      geom_bar(stat = "identity", fill = "#1a9988", alpha = 0.8, width = 0.4) +
      labs(title = paste("Count of", var)) +
      theme_minimal() +
      theme(
        legend.position = "none", 
        plot.title = element_text(size = 12),
        axis.title.x = element_blank(),
        axis.text.x = element_text(angle = 45, hjust = 1)
      )
    
    plot2 <- ggplot(mean_price_data, aes(x = !!sym(var), y = mean_log_price)) +
      geom_bar(stat = "identity", fill = "#eb5600", alpha = 0.8, width = 0.4) +
      labs(title = paste("Mean Log Price by", var)) +
      theme_minimal() +
      theme(
        legend.position = "none", 
        plot.title = element_text(size = 12),
        axis.title.x = element_blank(),
        axis.text.x = element_text(angle = 45, hjust = 1)
      ) 

    
    print(plot1 | plot2)
    
  }
```

### Selected variables

+ nhoods_name
+ landuse
+ LPSS_PER1000
+ HPSS_PER1000

```{r}
plot_list <- list()  # Store all plots
for (var in factor_vars) {
    count_data <- property_FE_lm_grouped %>%
      group_by(!!sym(var)) %>%
      summarise(count = n())
    
    mean_price_data <- property_FE_lm_grouped %>%
      group_by(!!sym(var)) %>%
      summarise(mean_log_price = mean(log_sale_price.x, na.rm = TRUE))
    
    plot <- ggplot(mean_price_data, aes(x = !!sym(var), y = mean_log_price)) +
      geom_bar(stat = "identity", fill = "#1a9988", alpha = 0.8, width = 0.3) +
      labs(title = paste("Mean Log Price by", var)) +
      theme_minimal() +
      theme(
        legend.position = "none", 
        plot.title = element_text(size = 12),
        axis.title.x = element_blank(),
        axis.text.x = element_text(angle = 45, hjust = 1)
      ) 

    
    plot_list[[var]] <- plot
}

# Arrange plots in a 2x2 grid
final_plot <- wrap_plots(plot_list) + plot_layout(ncol = 2)
print(final_plot)
```

## Summary of Features